### 2021-05-05 Allocation_tool/scripts/climate_change.Rmd
Included a mardonw with climate change text

---
output:
  bookdown::html_document2: 
    toc: yes
header-includes:
    - \usepackage[utf8]{inputenc}
    - \usepackage{wrapfig}
    - \usepackage{lipsum}
---

# Climate change and its effect on marine species

By 2017, greenhouse gas emissions from human activities are estimated to have warmed the Earth's surface by $1 \pm 0.2^ oC$ above pre-industrial levels causing changes in ocean conditions [@IPCC:2018ws]. The ocean has absorbed 93% of heat produced from greenhouse gas emissions since 1970 [@Rheim:2013fv; @IPCC:2019tn]. As a consequence, global sea surface temperature (SST) has increased by 0.63$^o$C relative to 1850-1900 [@IPCC:2019tn]. Also, an assessment by the Intergovernmental Panel on Climate Change (IPCC) has concluded that marine heatwaves have become more frequent [@Frolicher:2018dc], and ice sheets and glaciers have lost mass at an average rate of 220 $\pm$ 30 Gt $yr^{–1}$ between 2006 and 2015 [@IPCC:2019tn]. The ocean has also sequestered about 30% of the carbon dioxide ($CO_{2}$) emissions from the atmosphere since the late 1980's, causing a pH decrease of 0.1 unit that corresponds to a 26% increase in acidity [@Ross:2011kd; @IPCC:2019tn]. Moreover, increasing stratification of oceanic waters, changes in ventilation and biogeochemistry have led to a 0.5 to 3.3% loss of oxygen content in the upper layer (1000 m) of the open ocean between 1970 and 2010. During the same period, oxygen minimum zones have expanded by 3 to 8% in volume [@IPCC:2019tn].

![Past and future changes in sea surface temperature, surface ocean pH and oxygen content (100-600 m depth). Adapted from IPCC (2019)](../www/ipcc.png)

There are substantial regional variations in the $CO_{2}$-induced changes in ocean properties. For example, the Southern Ocean accounted for 35 to 43% of the total heat gain in the upper 2000 m layer of the global ocean between 1970 and 2017, and 25% of the global oxygen decline between 1970 and 1992 [@IPCC:2019tn]. Large reductions in oxygen have also been recorded for the north Pacific [@IPCC:2019tn]. In terms of pH, the largest declines between 1991 and 2011 were in the Indian Ocean (--0.027 units decade$^{ – 1}$), eastern Equatorial Pacific (--0.026 units decade$^{–1}$) and the South Pacific subtropical oceans (--0.022 units decade$^{–1}$) [@IPCC:2019tn].

Ocean properties are expected to continue to change in the $21^{st}$ century, although the intensity of these changes depends on the decisions' society will take in the future in relation to carbon mitigation [@Gattuso:2015jz; @IPCC:2019tn]. The IPCC assessed the future climate using scenarios of greenhouse gas concentration in the atmosphere. These scenarios, known as Representative Concentration Pathways (RCPs), range from a "strong climate mitigation" (RCP 2.6) scenario to a "no mitigation" scenario where society has no effective greenhouse gases mitigation policies (RCP 8.5). Under RCP 2.6, radiative forcing increases 2.6 $W/m^2$ by 2100 relative to pre-industrial conditions while RCP 8.5 leads to a radiative forcing of 8.5 $W/m^2$ [@vanVuuren:2011cs]. Under RCP 2.6, SST is projected to increase by 1.6$^o$C (CI[^1]: 1.1 - 2.0$^o$C) by 2031 - 2050 relative to 1850-1900, maintaining this trend (1.6$^o$C; CI: 0.9 - 2.4$^o$C) towards the end of the $21^{st}$ century. In contrast, following RCP 8.5 would result in an increase of 2$^o$C (CI: 1.5 - 2.4$^o$C) by 2031 - 2050 reaching 4.3$^oC$ (CI: 3.2 - 5.4$^o$C) by the end of the $21^{st}$ century relative to the same time period [@IPCC:2019tn; @Riahi:2011dk]. In addition, the IPCC considers two intermediate scenarios, RCPs 4.5 and 6.0. While RCP 4.5 leads to a radiative forcing of 4.5 $W/m^2$ by 2100 relative to pre-industrial conditions and RCP 6.0 leads to 6.0 $W/m^2$ [@Thomson:2011iy; @Masui:2011gr]. <!-- , RCP 1.5 leads to a much lower radiative forcing, 1.9 $W/m^2$. The RCP 1.5 is a scenario that align most closely to aspired carbon mitigation level under the Paris Agreement of the United Nations Framework Convention on Climate Change [@IPCC:2019tn; @IPCC:2014bz].  --> Here, I focus on RCP 2.6 and RCP 8.5 to capture the lower and upper range of climate change impacts on shared fish stocks and fisheries.

[^1]: In this case, the confidence interval used by the IPCC lies between the 17--83% probability range.

The ocean is projected to become warmer, less oxygenated and with lower pH in the $21^{st}$ century relative to the present day (e.g., 1986-2005) under all RCPs [@IPCC:2019tn]. Specifically, by 2050, global mean SST is expected to increase by 0.65$^o$C (CI: 0.33 to 0.96$^o$C) under RCP 2.6 and 0.95$^o$C (CI: 0.60 to 1.29$^o$C) under RCP 8.5. The largest SST warming will be in the North Pacific, the tropical East Pacific, and in parts of the Arctic [@Gattuso:2015jz]. At the same time, surface pH will be reduced by -0.072 (CI: --0.072 to --0.072) units under RCP 2.6 and 0.108 (CI: --0.106 to --0.110) units under RCP 8.5, with the Arctic experiencing the largest decrease [@IPCC:2019tn; @Gattuso:2015jz]. Dissolved oxygen within the first 600 m will also be reduced by -0.9% (CI: -0,3 to -1.5) under RCP 2.6 and -1.4% (CI: --1.0 to --1.8) under RCP 8.5 [@IPCC:2019tn]. The subsurface layer of mid-latitude regions is prone to see the largest changes in dissolved oxygen [@Gattuso:2015jz]. By the end of the $21^{st}$ century under RCP 8.5, ocean surface will be 3.2$^o$C warmer, --0.4 pH units lower, and have 3.9% less dissolved oxygen relative to the pre-industrial level. In contrast, under RCP 2.6, ocean surface will be 1.2$^o$C warmer, pH will be --0.14 units lower and oxygen will be about 1% less than the pre-industrial level [@IPCC:2019tn; @Gattuso:2015jz].

In responses to the change in ocean properties, marine species are altering their productivity, distribution, and phenology [@Scheffers:2016gz]. The type and magnitude of the responses will vary depending on the oceanic region, taxonomic group and life history and ecological characteristics of the organisms [@Poloczanska:2016kk; @Gattuso:2015jz; @Scheffers:2016gz]. Changes in maximum sustainable yield (MSY) of 235 populations of 162 species have been related to increasing water temperature. Between 1930 and 2010, MSY decreased by 4.1% (1.4 million tons) with some ecoregions (e.g., East Asia) experiencing losses of up to 35% [@Free:2019jq]. Future projections suggest that by the end of the $21^{st}$ century, 41 and 91% of global fish stocks will see a decline in MSY relative to 2012 under RCP 2.5 and 8.5 respectively [@Gaines:2018sg]. Moreover, shifting distributions has been the most widely documented response of marine species to ocean warming [@Poloczanska:2016kk; @Gattuso:2015jz]. Marine species, from phytoplankton to top predators have shifted their distribution ranges, mostly poleward, by an average of 72.0 $\pm$ 13.5 km per decade [@Poloczanska:2016kk]. These shifts are consistent with projections from species distribution models and are expected to continue in the $21^{st}$ century, even under a low emission scenario (RCP 2.6) [@Cheung:2010dt; @Barange:2014kr; @GarciaMolinos:2015fk]. Warming and less oxygenated waters have also been related to changes in size of marine fishes [@Oke:2020hi; @Forster:2012cq; @Audzijonyte:2020hy]. Overall, fish body size is expected to decrease under a warming ocean [@Cheung:2012fo; @Pauly:2017hpa; @Pauly:2018hq] with variations between regions and species [@Audzijonyte:2020hy].

Shifts in exploited marine species abundance, distribution and phenology will consequently impact the economics of fisheries [@Lam:2016dy; @Sumaila:2019ew; @Sumaila:2019gh] and food security [@Bell:2017hm; @Golden:2016gd] of communities that depend on them. It was estimated that in 2010, global fish landings were about 100 million tonnes (most of it destined to direct human consumption) generating about USD 150 billion [@Tai2017]. Estimates under RCP 8.5 suggest that global catches and revenue from fisheries could drop by 8 and 10%, respectively by 2050 with large regional differences [@Lam:2016dy]. For example, under RCP 8.5, the Exclusive Economic Zone (EEZ) of many tropical countries are expected to see \>50% reductions in both catch and revenue [@Lam:2016dy]. This is the case of many Pacific Islands nations like the Philippines where fish represents a high proportion of animal-sourced food [@Asch:2018ca; @FAO:2018KK; @Bell:2017hm]. On the contrary, high latitude EEZs such as the Russian in the north and the Chilean in the south could see increases of up to 30% [@Lam:2016dy]. Similarly, societal vulnerability to climate change impacts on fisheries will not be the same across coastal communities. For example, indigenous fishing communities that rely on marine resources for food security and wellbeing are particularly vulnerable to climate change due to reduced access to resources [@Weatherdon:2016ws; @CisnerosMontemayor:2016gq]. Reducing greenhouse gas emissions to limit warming to a 1.5$^o$C increase by the end of the $21^{st}$ century (RCP 2.6) could increase fisheries catch potential by up to 30% [@Cheung:2016tt]. This would result in a USD 4.6 billion annual revenue increase benefiting 75% of coastal nations from which \~90% represent developing countries [@Sumaila:2019ew].

<!-- ## Transboundary fisheries management under climate change -->

<!-- Current transboundary fisheries management might not be adapted to cope with shifts in species distributions under climate change [@Pinsky:2018cb, Chapter 4]. When a fish stock shifts from one EEZ to another due to climate change, fisheries in the states that lose the resource may opt to maximize the exploitation of the fish stock. In contrast, the states that benefits from an increasing share of the fish stock may have more incentive to conserve it because of its increasing stake on the state's fisheries resources [@Diekert:2017fa]. As climate change continues to drive shifts in biogeography of exploited species, fish stocks are expected to expand into EEZs they did not occupy before. In some cases, abundance will decrease in the lower latitudinal range of the species potentially exacerbating conflicts over stock's ownership and quota allocation [@Pinsky:2018cb; @Oremus:2020id]. Despite historical and projected changes in species distributions, today, many treaties concerning on the management of shared stocks are not well equipped to respond to stock shifts [@CullisSuzuki:2010fi; @Pinsky:2018cb; @Sumby:2021gw, Chapter 4]. Specifically, some basic fisheries management strategies, such as quota allocations, are based on historical catch proportions, or stock's historical distribution, and could become outdated as stocks shift to cope with a changing ocean [@Pinsky:2014by; @Beaugrand:2011fd, Chapter 4]. Lack of or insufficient adaptation of management of shared stocks to changing ocean conditions and stocks distributions will impact their effectiveness in ensuring the sustainability and benefits from the fisheries [@Sumaila:2020tp]. -->

<!-- Both the magnitude and pattern of range shifts are important to characterize the risk and impacts of shifting shared stocks on transboundary fisheries management [@Pinsky:2012kq; @Link:2010ei]. The distribution of a fish stock can change in multiple ways. For example, the stock range can shift from one EEZ to another while approximately maintaining its total range size. Also, the distribution of the stock can contract from or expand into a neighboring EEZ, resulting in a decrease or increase in range size, respectively (Figure \@ref(fig:ShiftsFigure)). The magnitude and characteristics of range shifts of shared stocks will have different consequences to their management [@Link:2010ei]. Historically, such shifts have caused conflict between neighboring nations and compromised the sustainability of diverse shared stocks. Prominent examples include the 1980's conflict over Pacific salmon between the United States (US) and Canada [@Miller:2004fpa; @Song:2017va]. -->

<!-- In addition to the magnitude and pattern of range shifts, it is important to understand the time frame at when the shifts and their impacts on fisheries will become apparent [@Pinsky:2014by; @Hawkins:2012hc]. The IPCC Special Report on the Ocean and Cryosphere in a Changing Climate used the concept of "time of emergence" (ToE) to discuss the challenge of changing ocean conditions, marine ecosystems and their challenges to ocean governance [@IPCC:2019tn]. As defined by the IPCC (adopted from Hawkins and Sutton, 2012), ToE is the moment in time when an anthropogenic change signal (e.g., future trend) rises above the background noise of natural variability (e.g., historical variation) [@IPCC:2019tn]. The premise behind ToE is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background of natural climate variability [@Hawkins:2012hc]. The concept of ToE has been extensively used to determine the time by which diverse oceanic variables (e.g., temperature, oxygen, $CO_2$) have raised above natural variability (e.g., emerged) in the global ocean [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Hawkins:2012hc; @Mahlstein:2011fy]. However, this concept has not been widely used on studies related to climate change and fisheries, until a recent study investigating the impacts of marine heatwaves in selected fisheries of the northeast Pacific [@Cheung:2020ja]. The concept ToE is useful to elucidate the impacts of climate change on transboundary fisheries and the consequences for their governance. For example, in the US, the arrival of the jumbo squid *Dosidicus gigas* to Washington state in 2009 created a new fishery under a *de facto* open access regime due to a lack of existing regulation. The time it took policy to be implemented was too long, and the fishery was over-exploited [@Pinsky:2014by]. Moreover, since 2007 when Atlantic mackerel (*Scomber scombrus*) extended its distribution to Iceland's EEZ, the European Union (EU) and Norway, and Iceland have been in dispute over fishing rights and quota allocation [@Spijkers:2017ij]. -->