### 2021-05-05 Allocation_tool/scripts/about.Rmd
Included about mrkd.

# Project background

Climate change is driving shifts in fish stock distributions that will affect resource availability across different jurisdictions around the world. Elinor Ostrom’s fundamental principles of common pool resource governance include a requirement for clear boundaries over resource units. However, when stock ranges shift, governance systems and boundaries are no longer aligned. This misalignment has already created conflicts around access and the re-distribution of benefits flowing from the stocks. This is a global challenge that is likely to increase as the effects of climate change intensify in the coming years. Specific problems that arise from the movement of stocks across jurisdictions include: overfishing; conflicts over access; unfairness to stakeholders who have borne the costs of resource stewardship but cannot capture the benefits due to stock movement; and high costs and conflicts associated with re-negotiating allocation, which is typically the most contentious aspect of fishery management.

# Objective and research questions

The objective of this tool is to evaluate policy options for managing fish stocks that are shifting across management boundaries, and to assess the socio-economic benefits and tradeoffs of these options for two important fisheries in the U.S. Mid-Atlantic.

1. Would  adaptive allocation  have reduced fuel use and carbon emissions, had it been applied when black sea bass and summer flounder stocks started to shift?

2. Would adaptive allocation have a positive economic impact on these fisheries, had it been applied when black sea bass and summer flounder stocks started to shift?

3.	How can conflicts around allocation be reduced in the context of shifting stocks?

4.	Where else in the U.S.  could this type of solution be applied to alleviate the challenges that arise when stocks shift?

# Partners

- [Arielle Levine, San Diego State University](https://geography.sdsu.edu/people/bios/levine)

- [Rod Fujita, Environmental Defense Fund](https://www.edf.org/people/rod-m-fujita)

- [Katie Longo, Marine Stewardship Council](https://www.researchgate.net/profile/Catherine-Longo)

- [Olaf Jensen, University of Wisconsin, Madison](https://limnology.wisc.edu/staff/jensen-olaf/)

- [Scott Crosson, NOAA](https://www.fisheries.noaa.gov/contact/scott-crosson-phd)

- [Chris Dumas, University of North Carolina Wilmington](https://csbapp.uncw.edu/data/fs/vita.aspx?id=8307)

- [Juliano Palacios-Abrantes, University of Wisconsin, Madison](https://limnology.wisc.edu/staff/palacios-abrantes-juliano/)

![](../www/logo_all2.png)