### 2021-08-13 Scripts/manuscript_acrossboundaries.Rmd
included results

---
title: "Challenges on quantifying fish range shifts across management boundaries"
subtitle: "*target journal* PlosOne"
output: word_document
---

<!-- # Journals -->
<!-- Ecography, ecology oriented -->
<!-- -	Conservation biology, -->
<!-- -	Fisheries journals; Fisheries -->

<!-- # Extra Ideas… -->

<!-- - Expanding from US to international -->
<!-- - Fishing communities and processing plans are fixed in time and space (e.g., Salmon in Canada, Peruvian anchoveta) -->

# Introduction

Fish range shifts have been occurring rapidly.

When harvested species shift across management boundaries, there can be a growing disconnect between the management jurisdictions in which fish are found and the management jurisdictions which hold harvest rights.

While the problem is increasingly well known, quantifying it is difficult for two reasons: (1) management jurisdiction in the ocean is often ill-defined, especially at the subnational level or beyond exclusive economic zones, and (2) quantifying fish distributions in relation to these boundaries is difficult, even where fishery-independent surveys exist, because of seasonal and interannual variation.

Here we examine both of these challenges using fish distributions relative to states in the Northeast U.S. as an example.

# Methods

We expanded the grid following two methods (Fig. 1).

![Figure 1. Spatial regions within the United States' North East Exclusive Economic Zone. A) Spatial buffers from each State's waters. B) Spatial buffers from the main fishing ports. States are aligned by average latitude from north to south. ME = Maine; NH, New Hampshire; MA, Massachusetts; RI, Rhode Island; CT, Connecticut; NY, New York; NY, New Jersey; DE, Delaware; MD, Maryland; VA Virginia; North Carolina.](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F1_buffer_figure.jpg)

# Results

![Figure 2. Change in the proportion of species distribution per State in recent years (2002 - 2014) relative to the reference years (1990-2001) following the fishing ports (top) and State waters (middle) approach. Bottom pannel showing the percentage difference between both approaches. ](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F2_proportion_chg_diff.jpg)

![Figure 3. Aggregated distribution proportion per state of Paralichthys dentatus, Stenotomus chrysops and Centropristis striata over time. Each color represent a State ordered latitudinal. Grey box represents the base-period used for the fisheries quota allocation ](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F3_area_plot_avg_agg_sptl.jpg)

![Figure 4. Proportion of the distribution of Paralichthys dentatus, Stenotomus chrysops and Centropristis striata over time. Each color represent a State ordered latitudinal. Grey box represents the base-period used for the fisheries quota allocation ](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F4_area_plot_avg_spp_sptl.jpg)