### 2021-10-05 Scripts/manuscript_acrossboundaries.Rmd
included alocation maps for spatial and season differences

---
title: "Quantifying fish range shifts across poorly defined management boundaries"
subtitle: "DRAFT. Target journal *PlosOne*"
author: "Palacios-Abrantes Juliano, Crosson Scott, Dumas Chris, Fujita Rod, Levine Arielle, Longo Catherine, Jensen Olaf"
output: 
  bookdown::word_document2:
    number_section: no
    toc: no
---

<!-- # Journals -->
<!-- Ecography, ecology oriented -->
<!-- -	Conservation biology, -->
<!-- -	Fisheries journals; Fisheries -->

<!-- # Extra Ideas… -->

<!-- - Expanding from U.S. to international -->
<!-- - Fishing communities and processing plans are fixed in time and space (e.g., Salmon in Canada, Peruvian anchoveta) -->

```{r setup, include=FALSE}

library(MyFunctions)
MyFunctions::my_lib(c("dplyr","tibble","knitr","kableExtra","png","grid"))

```

# Introduction

Range shifts are among the most documented responses of marine fauna to the effects of climate change in the ocean [@Poloczanska]. From zooplankton to bony fish to seabirds, marine fauna has been documented to expand their distribution up to 800 km per decade or to contract it for about 200 km per decade [@Poloczanska]. Such change is often due to marine species following a poleward temperature gradient in the search for colder waters, although, in some cases such species are following local environmental factors such as upwelling systems [@Clarke:2020], or moving to deeper colder waters [@Pinsky:2013]. Regardless of the shift direction, species distributions are projected to continue to happen in the incoming years [@Cheung:2010; @Barange2014; @IPCC2019], even if society engages in a strong mitigation climate change scenario (i.e., limiting increased global temperature to $1.5^o$) as proposed by the Paris Agreement [@Cheung:2016].

The change in species distribution will be particularly challenging for the management of stocks shared across management jurisdictions [@PalaciosAbrantes:2020sr]. When a stock falls within to two or more spatial regulatory regions, managers must often coordinate efforts to maintain sustainability of the stock [@Clark1979; @Munro2004]. However, there can be a growing disconnect between the region of the ocean where fish are found and the management jurisdictions which hold harvest rights [@PalaciosAbrantes:2020sr]. Such mismatch is of specific concern as species shift their known distribution due to climate change [@PalaciosAbrantes_2020pn; @oremus_2020, @pinsky_2018; @Baudron2020; @Miller2013]. Historical shifts in important fish stocks have resulted in both unsustainable harvest and international conflict as in the case of the United States and Canada over sockeye salmon (*Oncorhynchus nerka*) and the European Union and Ireland over mackerel (*Scomber scombrus*) [@miller2013; @spijkers2017].

In the United States (U.S.) North East (NE) Atlantic, many important fished species such as american lobster (*Homarus americanus*), yellowtail flounder (*Limanda ferruginea*), summer flounder (*Paralichthys dentatus*) and red hake (*Urophycis chuss*) have been shifting their distribution poleward  [@pinsky_fogarty_2012], and modelling exercises suggest that NE Atlantic species will shift will average over 600 km towards the end of the $21^{st}$ century under a high emission climate change scenario [@Morley:2018]. The mismatch between implemented management strategies and shifts in stocks across management jurisdictions compromises management and sustainability [@Baudron2020; @pinsky_fogarty_2012]. Fishers in the NE Atlantic have been adapting to such shifts by following the fish, in some cases, or changing target speices, in others [@polo_2021]. While this problem of shifting stocks across management jurisdictions is increasingly well known, quantifying it is difficult for two main reasons: (*i*) management jurisdictions in the ocean are often ill-defined, especially at the sub-national level or beyond Exclusive Economic Zones (EEZ), and (*ii*) quantifying fish distributions in relation to these boundaries is difficult, even where fishery-independent surveys exist, because of seasonal and interannual variation.

**[Short paragraph on the ill-defined boundary challeng - do we privilege a state's geographic claims or the inherent claims to the resource of existing fishing ports?]**

**[Short paragraph on the seasonality of fish distributions in temperate waters - when do we measure fish distributions?]**

Here, we examine both of these challenges by quantifying the historical distributions of three important fished species relative to 11 states of the Northeast Atlantic U.S that jointly manage these stocks. Specifically, we set two main objectives. First, we examine the difference between two different approaches to define spatial management jurisdictions within the U.S. Federal waters. Second, we examine differences in the species' historical distribution between Spring and Fall to examine seasonal variation in these stock's distributions.

 "How does the choice of survey timing and boundary extrapolation method influence our perception of fish distribution changes relative to management boundaries?"

# Methods

The current analysis focused on black sea bass (*Centropristis striata*), summer flounder (*Paralichthys dentatus*) and scup (*Stenotomus chrysops*), all species managed under the Mid-Atlantic Council Management Plans [(MAFMC)](https://www.fisheries.noaa.gov/new-england-mid-atlantic/commercial-fishing/new-england-mid-atlantic-fishery-management-plans) since 1996. The management unit extends from the U.S.-Canadian border to the southern border of North Carolina for summer flounder, and Cape Hatteras (North Carolina) for scup and black sea bass [MAFMC](https://www.mafmc.org/sf-s-bsb/). The fisheries are managed cooperatively by the States through the Commission in state waters (0-3 miles), and through the MAFMC and NOAA Fisheries in federal waters (3-200 nautical miles). An annual coast wide commercial quota is implemented and since 2003 further allocated to the States of Maine, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York, New Jersey, Delaware, Maryland, Virginia and North Carolina [@REFMACMP-memorandum] (Table \@ref(tab:QuotaTable)). [OJ](I don't this is right.  For summer flounder, at least, I thought there was a specific year that the quota allocation was based on.) The state-wide quota allocation is based on the historical landings from 1980-2001, approximately.

```{r QuotaTable, eval = F, echo = F}

tibble(
  "State" = c(
    "Maine (ME)",
    "New Hampshire (NH)",
    "Massachusetts (MA)",
    "Rhode Island (RI)",
    "Connecticut (CT)",
    "New York (NY)",
    "New Jersey (NJ)",
    "Delaware (DE)",
    "Maryland (MD)",
    "Virginia (VA)",
    "North Carolina (NC)"
  ),
    "Centropristis striata" = c(
      0.0040, #ME
      0.0040, #NH
      0.1564, #MA
      0.1323, #RI
      0.0367, #CT
      0.0857, #NY
      0.2010, #NJ
      0.0411, #DE
      0.0888, #MD
      0.1614, #VA
      0.0888 #NC
    ),
    "Paralichthys dentatus" = c(
      0.0004756, #ME 
      0.0000046, #NH 
      0.0682046, #MA 
      0.1568298, #RI 
      0.0225708, #CT 
      0.0764699, #NY 
      0.1672499, #NJ 
      0.0001779, #DE 
      0.0203910, #MD
      0.2131676, #VA
      0.2744584 #NC
    ),
    "Stenotomus chrysops" = c(
      0.0012101, #ME
      0.0000000, #HN
      0.2158729, #MA
      0.5619456, #RI
      0.0315399, #CT
      0.1582466, #NY
      0.0291667, #NJ
      0.0000000, #DE
      0.0001190, #MD
      0.0016502, #VA
      0.0002490 #NC
    )
) %>% 
  mutate_if(is.numeric,round,3) %>% 
  mutate_if(is.numeric,~.*100) %>% 
  # readr::write_csv("quota_table.csv")
  # group_by() %>% #just checking they add up
  # summarise_if(is.numeric,sum)
  kable("latex",
        caption = "State wide commercial quota allocation for black sea bass (*Centropristis striata*), summer flounder (*Paralichthys dentatus*) and scup (*Stenotomus chrysops*) according to the Mid-Atlantic Fishery Management Council. Values shown as percentage of total quota. States ordered latitudinally"
        )
```

Distribution data for all three species was gathered from NOAA's yearly Northeast Fisheries Science Center Spring and Fall Bottom Trawl Surveys [data](https://www.fisheries.noaa.gov/region/new-england-mid-atlantic#science) provided by [Ocean adapt](https://oceanadapt.rutgers.edu/) and standardized as catch per unit of effort (CPUE) [@Pinsky2013]. CPUE was defined as trawl-catch in weight per area swept (kg/ha). For the extrapolation and subsequent analysis we focused on the Atlantic U.S. EEZ spatial division using the *Sea Around Us* shape file (updated 2020; http://www.seaaroundus.org). The EEZ shapefile was gridded to a 3 x 3 degrees matrix. We extrapolated trawl CPUE data to the U.S. EEZ from the southern border of North Carolina to the U.S.-Canadian border by following a triangular irregular surface (TIS) method using the *R* package *interp* [@REFInterp]. Overall, the TIS method creates multiple non-overlapping triangles that cover the entire extrapolated region (i.e., the EEZ) from an initial set of sample points (i.e., NOAAs survey) generating a spatially-continuous surface [@REF; @REF]. The extrapolation was done separately for both Fall and Spring seasons covering the time period of 1971 to 2019.

## Determining regulatory units

Although the stock-wide quota for our three target species is allocated among the states, U.S. Federal waters are not sub-divided into state-owned fisheries management regions (**hereafter refereed as regulatory units**). This lack of regulatory units within federal waters creates a challenge for assessing a stock's distribution with respect to individual states. There are a potentially infinite number of ways to geographically determine regulatory units with the EEZ's 200 nautical miles. Here we focus on two methods based which likely bracket much of the range of options.

### State waters expansion

<!-- Office for Coastal Management, 2021: Federal and State Waters from 2010-06-15 to 2010-08-15. NOAA National Centers for Environmental Information, https://www.fisheries.noaa.gov/inport/item/54383. -->

The first approach consists on expanding the current geographical distribution of States and their statutory state waters. For this approach we adapted the state's waters boundaries determined by the Federal and State Waters shapefile created for the Ocean Reporting Tool in April, 2018 ([NOAA's Office for Coastal Management, 2021](https://catalog.data.gov/dataset/federal-and-state-waters)). This shapefile consists in multiple polygons each one representing the statutory 3 miles waters that each state owns (Figure \@ref{fig:figureS1}). Each state's polygon  from Maine to Virginia was expanded to the 200 nm EEZ boundary using the *st* package in *R* [@REF]. The resulting polygons were also gridded into a 3 x 3 deg. matrix. The regions where the grid cells overlapped were equally allocated to the sharing states.

### Fishing ports approach

The second approach to determine regulatory units gives greater weight to the historical landings of fishing ports. For this approach we used NOAA Fisheries landings-by-port data [@REF_NOAA]. Landings from each species and port were averaged from 2010 to 2019 noting that 2020-2021 were removed to avoid using COVID19-impacted years. We then estimated the total port-wide landings of the averaged period and the proportion that each port had in terms of landings *per* species. We selected the main ports landing each species, defined as those concentrating the top $75{^th}$ percentile of landings. Note that, while species often shared top ports, their final number and location varied according to each species (Figure \@ref{fig:figureS1}). Moreover, we tested the sensitivity of our results to the arbitrary $75{^th}$ percentile threshold and found that... (*to be done*). Each port represented a point in the map (i.e., latitude and longitude degrees) whose coordinates were manually collected using [googlemaps](www.googlemaps.com). We then drew a polygon from each port to the EEZ and gridded the expanded polygon following the same method as the *Fishing ports approach*. The expanded polygon of each port was associated to the state where the port is located and overlapping grids were equally allocated to the sharing states.

## Estimating each state's CPUE

We overlay the specie's interpolation with the regulatory units and estimated the distribution and CPUE proportion that each state had under both approaches. The interpolated CPUE was divided equally in cases where two or more states had the same grid. For example, if New York and New Jersey both shared one grid, then the proportion of that grid would be allocated 50% to each state, alternatively, if Connecticut was also sharing the grid, then the proportion would be 33.3% to each state. Note that this step does not affect the interpolation method but the way the CPUE is allocated by state.

Finally, we estimated the CPUE percentage change today relative to the historic period. For that matter we define the historic period as 1980 to 2000 representing the state allocations of the commercial fisheries originally implemented in 2003 [@RefDocument], and “today” as the average of 2009 to 2019 representing the more recent data available. The parentage changes between the historic and present CPUE ($\Delta{B}$) was estimate for each of the two regulatory units approaches as;

$$ \Delta \alpha=\frac{(\alpha_t-\alpha_h)}{|\alpha_h|}*100$$

Where $\alpha_t$ is the CPUE today (average of 2009 to 2019) and $\alpha_h$ is the CPUE at the historical period (average of 1980 to 2000). In addition, we estimated the differences in ($\Delta{\alpha}$) between survey seasons ($\Delta{\alpha}_{seas}$) and between regulatory units ($\Delta{\alpha}_{reg}$) approach by subtracting from each other:

$$ \Delta \alpha_{seas} = |\Delta \alpha_{fp}-\Delta \alpha_{sw}| $$
$$ \Delta \alpha_{reg} = |\Delta \alpha_{f}-\Delta \alpha_{s}| $$

Where $\Delta{\alpha}_{fp}$, $\Delta{\alpha}_{sw}$, $\Delta{\alpha}_{f}$, and $\Delta{\alpha}_{s}$ represent the CPUE percentage change today relative to historical value for fishing ports, state waters, Fall and Summer approaches, respectively.

All of the analysis was done on R-Studio statistical software and *R* version 4.1.0 (*Camp Pontanezen*) [@REF_Rstudio] and associated packages (Table S1). All code can be found in https://github.com/jepa/AcrossBoundaries.
<!-- *ggmap*, *sf* [@REF], *st* [@REF], *tidyverse* [@REF], *tools* [@REF], *readr* [@REF], *data.table* [@REF], *maps* [@REF], *viridis* [@REF], *wesanderson* [@REF], *knitr* [@REF], *kableExtra* [@REF], *plotly* [@REF], *ggrepel* [@REF], *ggsflabel* [@REF], *janitor* [@REF], *interp* [@REF], *rnaturalearth* [@REF], and *readxl* [@REF].  -->

# Results

The choice of boundary expansion method (e.g., fishing ports or state waters) is highly consequential for determining the spatial extent of each state's regulatory unit (Figure \@ref{fig:bufferFig}). For example, using the state water approach, Delaware and Maryland have regulatory units of *XX* and *YY* hectares, respectively. Because neither state has a port with commercial landings of black sea bass, scup, or summer flounder that exceeds the $75^{th}$ threshold, their regulatory units based on the fishing ports approach include only their state waters out to 3 nm. In contrast, New Jersey's regulatory unit is *X%* larger under the fishing port approach because New Jersey has *XX* ports meeting the $75^{th}$ landings threshold, and they are separated by *XX* km.

```{r bufferFig, eval = F, echo = T, fig.width = 3, fig.height = 2, echo=FALSE, fig.cap="Spatial regions within the United States' North East Exclusive Economic Zone. A) Spatial buffers from each state's waters. B) Spatial buffers from the main fishing ports. States are aligned by average latitude from north to south. ME, Maine; NH, New Hampshire; MA, Massachusetts; RI, Rhode Island; CT, Connecticut; NY, New York; NY, New Jersey; DE, Delaware; MD, Maryland; VA Virginia; North Carolina."}

# Call the image
img <- readPNG("/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F1_buffer_figure.png")
# Print the image
grid.raster(img)
```

![Figure 1. Spatial regions within the United States' North East Exclusive Economic Zone. A) Spatial buffers from each state's waters. B) Spatial buffers from the main fishing ports. States are aligned by average latitude from north to south. ME, Maine; NH, New Hampshire; MA, Massachusetts; RI, Rhode Island; CT, Connecticut; NY, New York; NY, New Jersey; DE, Delaware; MD, Maryland; VA Virginia; North Carolina. ](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F1_buffer_figure.jpg)

Seasonality plays an important role in determine the distributional range of black sea bass, scup, and summer flounder (Figure \@ref{fig:figInterpolation}). Latitudinal variations in distributions are notable for scup, whose historical northern limit goes from Massachusetts (latitude ~42.4) in spring to the northern coast of Maine (latitude ~44) in the fall. Moreover both scup and black sea bass have expanded their distribution northward. For example, black sea bass Northern limit in spring expanded from New Jersey (latitude ~$40.6^o$) in the late 1980s to the border of Maine and New Hampshire (latitude ~$43^o$) in recent years. In addition, black sea bass centroid (i.e., the region with the highest CPUE) shifted $2.1$ degrees north since 1980, more than any of the three species. In the case of summer flounder, however, such latitudinal seasonality is less evident. Instead, the interpolation results suggests a longitudinal (or depth) expansion in the species range from a largely coastal concentration in the fall to an eastward shift in the spring. Such pattern seems to become even more pronounced in current spring conditions, relative to historical levels (Figure \@ref{fig:figInterpolation}).

```{r figInterpolation, eval = F, echo = T, fig.width = 3, fig.height = 2, echo=FALSE, fig.cap="Interpolation results for all species and seasons. Scale showing Catch per Unit of Effort (CPUE) at a log 10 scale. White cross represents the species' centroid defined as the region with the largest CPUE. Grey regions represent no CPUE"}

# Call the image
img <- readPNG("/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F2_Interpolation.jpg")
# Print the image
grid.raster(img)
```

![Figure 2. Interpolation results for all species and seasons. Scale showing Catch per Unit of Effort (CPUE) at a log 10 scale. White cross represents the species' centroid defined as the region with the largest CPUE. Grey regions represent no CPUE](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F2_Interpolation.jpg)

We overlay the developed regulatory units with the interpolated CPUE of black sea bass, summer flounder and scup to estimate each state's proportion of the total CPUE (Figures \@ref{fig:figSpatialDiff}). Under the state waters approach, each state has a regulatory unit and thus captures a part of the CPUE of each species. However, under the fishing ports approach some states end up without regulatory units resulting in no CPUE. This is the case of Maine, New Hampshire and Connecticut. On the contrary, New Jersey and Virginia end up with the largest benefits of the fishing ports approach. There are very few cases where the spatial approach does not affect the CPUE proportion of each state (Delaware, Massachusetts and North Carolina for black sea bass, North Carolina for summer flounder and Maryland for scup).

```{r figSpatialDiff, eval = F, echo = T, fig.width = 3, fig.height = 2, echo=FALSE, fig.cap = "Proportion of the Catch per Unit of Effort (CPUE) for each species by state undert two time periods. Top and middle rows represent the fishing port and state waters approach, respectiveley. Bottom row is the difference between each approach. All values in percentage"}

# Call the image
img <- readPNG("/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F3_proportion_chg.jpg")
# Print the image
grid.raster(img)
```

![Figure 3. Proportion of the Catch per Unit of Effort (CPUE) for each species by state undert two time periods. Top and middle rows represent the fishing port and state waters approach, respectiveley. Bottom row is the difference between each approach. All values in percentage. Differences in absolute values](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F3_proportion_chg.jpg)

Seasonal differences are also evident, although smaller than regulatory unit differences, when estimate each state's CPUE proportion for all species (Figures \@ref{fig:figSeasonlDiff}). Here, half of the cases represent 1% or no seasonality difference like Connecticut's CPUE proportion of summer flounder. The largest difference are in the latitudinal extremes with Maine currently getting 4% more CPUE in the fall than in spring as opposite to Virginia who gets 4% more in the spring than in the fall.

```{r figSeasonlDiff, eval = F, echo = T, fig.width = 3, fig.height = 2, echo=FALSE, fig.cap = "Proportion of the Catch per Unit of Effort (CPUE) for each species by state undert two time periods. Top and middle rows represent the fishing port and state waters approach, respectiveley. Bottom row is the difference between each approach. All values in percentage"}

# Call the image
img <- readPNG("/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/proportion_chg_seas.jpg")
# Print the image
grid.raster(img)
```

![Figure 4. Historical Catch Per Unit of Efort (CPUE) proportion of each species by season. Proportion presented in 10 years running mean percentage. States presented latitudinally from north to south. Differences in absolute values. ](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/proportion_chg_seas.jpg)

The proportion of the CPUE has increased for northern states

```{r figAreaPlot, eval = F, echo = T, fig.width = 3, fig.height = 2, echo=FALSE, fig.cap = "Proportion of the Catch per Unit of Effort (CPUE) for each species by state undert two time periods. Top and middle rows represent the fishing port and state waters approach, respectiveley. Bottom row is the difference between each approach. All values in percentage"}

# Call the image
img <- readPNG("/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F4_area_plot_avg_spp_sptl.jpg")
# Print the image
grid.raster(img)
```

![Figure 5. Historical Catch Per Unit of Efort (CPUE) proportion of each species by spatial approach. Proportion presented in 10 years running mean percentage. States presented latitudinally from north to south ](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/F4_area_plot_avg_spp_sptl.jpg)

# Discussion

# Supplements

```{r figsone, eval = T, echo = F}

```