### 2021-11-29 Scripts/spatial_analisis.Rmd
Updated F3 with plys/minus scake

---
title: "Spatial Analysis"
author: "Juliano Palacios Abrantes"
output:
  html_document:
  # html_notebook:
    fig_width: 6
    fig_height: 4
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float:
      collapsed: false
    theme: darkly
---

```{r setup, include=FALSE}
library(MyFunctions)
MyFunctions::my_lib(c("ggmap","sf","st","tidyverse","tools","readr","data.table","maps","viridis","wesanderson","knitr","kableExtra","plotly","ggrepel","ggsflabel","janitor","interp","rnaturalearth","readxl","cowplot"))

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

# Create specific state pallet from our friend Wes Andersson
state_pallet <- c(wes_palette(n = 5, name = "Darjeeling1"),
                               wes_palette(n = 5, name = "Darjeeling2"),
                               wes_palette(n = 3, name = "Royal1")
                  )

```

# Methods

## Spatial data preparation

### Load data

#### Maps data

We use three different shapefiles for the continental U.S. land mass, the State waters of maine, new hampshire, massachusets, connecticut, rhode island, new york, new jersey, delaware, maryland, virginia, north carolina and the North East EEZ. Moreover, we use the NOAA dataset of landings per port.

##### 1.- Land shapefile

Covers the US land territory for visualization. Data provided from the `map` package.

```{r land_sf, eval = T, echo = T, results = 'hide'}

States <- c("maine", "new hampshire", "massachusetts", "connecticut", "rhode island", "new york", "new jersey", "delaware", "maryland", "virginia", "north carolina")

# -------------- #
# US State Map (land)
# -------------- #

land_sf <- st_as_sf(map("state", plot = FALSE, fill = TRUE)) %>% 
  filter(ID %in% States) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA"),
         state = str_to_sentence(ID)
         )

# Visual exploration (o.k! )
ggplot(land_sf) +
  geom_sf(aes(fill = state)) +
  scale_fill_manual(values = state_pallet) +
  theme_dark()

```

#####  2.- EEZ shapefile

Used the *Sea Around Us* shapefle updated to June 2016.

```{r eez_sf, eval = T, echo = T}

# US EEZ map
path_world <- my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile", name = "SAUEEZ_July2015.shp")

# Load it!
eez_sf <- st_read(path_world) %>%
  rename(eez_name = Name) %>%
  st_transform(crs = 4326) %>% # 4326
  filter(eez_name == "USA (East Coast)") %>%
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) #0.1 for paper

# Visual exploration (o.k! )
ggplot(eez_sf) +
  geom_sf() +
  theme_dark()

```

#####  3.-  State waters

Covers the state waters of the NE US states. Data from [data.gov](https://catalog.data.gov/dataset/federal-and-state-waters).

*License: No license information was provided. If this work was prepared by an officer or employee of the United States government as part of that person's official duties it is considered a U.S. Government Work.*

```{r state_sf, eval = T, echo = T}

state_sf <-  st_read("~/OneDrive - UW-Madison/FederalAndStateWaters/FederalAndStateWaters.gdb") %>%
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  mutate(state = str_to_sentence(jurisdicti)) %>% 
  st_transform(crs = 4326) %>%   # for matching projections
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) #0.1 for paper

# Visual exploration (o.k! )
ggplot(state_sf) +
  geom_sf(aes(fill = state)) +
  scale_fill_manual(values = state_pallet) +
  theme_dark()

```

##### 4.- Fishing ports

###### 4.1 NOAA's data

Dataset provider: Joe Myers Joseph.Myers@accsp.org NOAA

Non-confidential dataset containing annual landings/value summaries by port for black sea bass, scup and summer flounder, from NC-ME for 1970-2020.

In order to make the report non-confidential, I am displaying as much of the port-level summary data as I can – and for cases where the port-level was confidential, I have first attempted to roll-up any confidential port-level summaries into state-level summaries within the same year.  If those state-level summaries continue to be confidential, I have further rolled-up those confidential ports across states into a coast-level summary.  Even at the coast level, there were several cases where not enough confidential ports could be combined to produce a non-confidential summary – and so for those rows it was necessary to redact the final landings/value data.  Fortunately, this only amounts to less than 0.5% of the total landings data (characterized in the “Pivot by Confidentiality” tab of the spreadsheet).

*License???.*

```{r}
# NOAA DATA

ports_path <- my_path("D","NOAA", name = "PalaciosJ - BSB, SFlounder, Scup by Port 1970-2020 - 2021.10.25.xlsx")

# Load ports
ports_data <- read_excel(ports_path,sheet = 2) %>% 
  clean_names() %>%
  mutate_if(is.character,
            ~ str_to_lower(.)
  ) %>% 
  mutate(
    species_name = ifelse(str_detect(common_name,"black"),"black sea bass",
                 ifelse(str_detect(common_name,"summer"),"summer flounder",
                        common_name)
    ),
    total_live_pounds = as.numeric(total_live_pounds),
    total_value = as.numeric(total_value),
    # Allocate all bsb NC data to Wanchese
    port_name = ifelse(species_name == "black sea bass" & state_postal == "nc" & port_name == "unknown","wanchese",port_name)
  ) %>% 
  filter(
    !port_name %in% c("**roll-up to state-wide**","**roll-up to coast-wide**","unknown")
  )
```

###### 4.1.1 Top ports

```{r}
# Get number of ports in the NOAA database
overall_port_data <- ports_data %>%
  filter(year >= 2010 & year <= 2019) %>% 
  group_by(species_name,port_name,state_postal) %>% 
  summarise(
    year_ton = sum(total_live_pounds),
    year_value = sum(total_value)
  )

# quantiles_df <- overall_port_data %>% 
#   group_by(species_name) %>% 
#   summarise(
#     "95th" = quantile(year_ton, 0.95),
#     "75th" = quantile(year_ton, 0.75),
#     "50th" = quantile(year_ton, 0.50)) %>% 
#   gather("tresh","quant",`95th`:`50th`)
# 
# 
# top_portsb <- overall_port_data %>% 
#   left_join(quantiles_df,
#             by = "species_name") %>% 
#   mutate(
#     treshold = ifelse(year_ton >= quant,tresh,NA) 
#   )

top_ports <- bind_rows(
  # 75th percentile used in project
  overall_port_data %>% 
    group_by(species_name) %>% 
    top_frac(0.25,year_ton) %>%  # top 75
    mutate(tresh = "75th"),
  # -------------------- #
  # For sensitivity values
  # -------------------- #
  # 95%
  overall_port_data %>% 
    group_by(species_name) %>% 
    top_frac(0.05,year_ton) %>%  
    mutate(tresh = "95th"), # top 95
  # 85%
  overall_port_data %>% 
    group_by(species_name) %>% 
    top_frac(0.15,year_ton) %>% 
    mutate(tresh = "85th"), # top 85
  # 65%
  overall_port_data %>% 
    group_by(species_name) %>% 
    top_frac(0.35,year_ton) %>%  # top 75
    mutate(tresh = "65th"),
  # 50%
  overall_port_data %>% 
    group_by(species_name) %>% 
    top_frac(0.5,year_ton) %>%  # top 50
    mutate(tresh = "50th")
)

```

###### 4.1.2 Ports coordinates

```{r sarahs_word_data}

noaa_ports <- data.table(
  port =str_to_sentence(c(
    "POINT JUDITH, RI", #41.359907914553574,-71.48323250872187,
    "BEAUFORT, NC", #34.7138879654054,-76.66239630117921,
    "HAMPTON, VA", #36.98600366773327, -76.35038306859545,
    "POINT PLEASANT, NJ", #40.08351847675102,-73.99190094110679,
    "NEWPORT NEWS, VA", #37.00933919295183, -76.48565862613168,
    "BELFORD, NJ", # 40.43918784734024, -74.0763054899462
    "MONTAUK, NY", #41.07342262868038, -71.93585836088813
    "HOBUCKEN, NC", #34.94844567982384, -76.74333419560425
    "WANCHESE, NC", #35.83524984550334, -75.61121671722921
    "NEW BEDFORD, MA", # 41.630494052326306, -70.91580669134055
    "CAPE MAY, NJ", # 38.94979719126508, -74.90280944790463
    "ORIENTAL, NC", # 35.013571084373126, -76.66793114333704
    "CHINCOTEAGUE, VA", # 37.91788094070204, -75.35947580478219
    "ENGELHARD, NC", # 35.262929243118776, -76.08680452670508
    "STONINGTON, CT", # 41.327282204484035, -71.9223721359251
    "LONG BEACH/BARNEGAT LIGHT, NJ", #39.752855929900484, -74.11329310115902
    "OCEAN CITY harbor, MD", #38.3802922004859, -75.05089736920318
    # Manually included these ports as top State ports from NOAA
    "Indian river, DE",
    "Portland, ME", #43.66344794728992, -70.22937554378885
    "Portsmouth, NH", #43.05334694203953, -70.7126760833714
    # More ports
    "Chincoteague, VA", # 37.94295896302465, -75.38415668579887
    "sea isle city,NJ ", # 39.14912656323854, -74.68797202847871
    "lewes, DE, ", # 38.78943640990542, -75.06343692610851
    "virginia beach, VA", # 36.859343385722504, -75.94328398140065
    "hampton bays, ny", #40.84229040189204, -72.50987544040999
    "mattituck, ny", # 41.05342836192293, -72.59899214258269
    "little compton, ri", # 41.4765103461213, -71.19474641579107
    "new london, CT", #41.341577207467914, -72.0829697106489
    "amagansett, ny", # 40.96173008633569, -72.11879306740798
    "barnegat light, nj", #39.75291152662383, -74.09927831868308
    "chatham (census name for chatham center), MA", # 41.658695421122474, -69.95466452556448
    "dennis, ma", #41.65121037263873, -70.13193663798509
    "falmouth (census name for falmouth center), ma", #41.53956327146526, -70.60001340709665,
    "freeport, ny", #40.62242387473109, -73.59162509402208
    "nantucket (census name for nantucket center), MA", #41.26616088210985, -69.95617277257985
    "newport, RI",
    "point lookout, ny", #40.587646134955875, -73.5763748512927
    "shinnecock indian reservation, ny", #40.865873983968754, -72.4420929172018
    "westport, ma"
  )),
  lat = c(
    41.359907914553574,
    34.7138879654054,
    36.98600366773327, 
    40.08351847675102,
    37.00933919295183, 
    40.43918784734024, 
    41.07342262868038,
    34.94844567982384, 
    35.83524984550334,
    41.630494052326306,
    38.94979719126508,
    35.013571084373126,
    37.91788094070204, 
    35.262929243118776, 
    41.327282204484035, 
    39.752855929900484, 
    38.3802922004859,
    38.62113593403746,
    43.66344794728992,
    43.05334694203953,
    37.94295896302465,
    39.14912656323854,
    38.78943640990542,
    36.859343385722504,
    40.84229040189204,
    41.05342836192293,
    41.4765103461213, 
    41.341577207467914,
    40.96173008633569, 
    39.75291152662383, 
    41.658695421122474,
    41.65121037263873, 
    41.53956327146526, 
    40.62242387473109,
    41.26616088210985, 
    41.485514346683956,
    40.587646134955875,
    40.865873983968754,
    41.49267875746873
  ),
  lon = c(
    -71.48323250872187,
    -76.66239630117921,
    -76.35038306859545,
    -73.99190094110679,
    -76.48565862613168,
    -74.0763054899462,
    -71.93585836088813,
    -76.74333419560425,
    -75.61121671722921,
    -70.91580669134055,
    -74.90280944790463,
    -76.66793114333704,
    -75.35947580478219,
    -76.08680452670508,
    -71.9223721359251,
    -74.11329310115902,
    -75.05089736920318,
    -75.06397021718345,
    -70.22937554378885,
    -70.7126760833714,
    -75.38415668579887,
    -74.68797202847871,
    -75.06343692610851,
    -75.94328398140065,
    -72.50987544040999,
    -72.59899214258269,
    -71.19474641579107, 
    -72.0829697106489,
    -72.11879306740798,
    -74.09927831868308,
     -69.95466452556448,
    -70.13193663798509,
    -70.60001340709665,
     -73.59162509402208,
    -69.95617277257985,
     -71.32263326975122,
     -73.5763748512927,
     -72.4420929172018,
    -71.05551342392039
  )
) %>% 
  separate(port, into = c("port_name","state_postal"), sep = ',') %>% 
  mutate(port_name = str_to_lower(port_name),
         state_postal = str_squish(state_postal))

# Include coordinates
noaa_ports_complete <- top_ports %>% 
  left_join(noaa_ports,
             by = c("port_name", "state_postal")) %>% 
  mutate(
    state_postal = str_to_upper(state_postal)
  )

# Check for missing ports
noaa_ports_complete %>% 
  filter(tresh == "75th",
    is.na(lat)) %>% 
  group_by(port_name,state_postal) %>% 
  summarise(n()) %>% 
  View()

# Save data for future analysis
write_csv(noaa_ports_complete,
          my_path("D","Partial", name = "top_ports.csv")
          )

noaa_ports_complete

```

###### 4.1.1 Visualization

```{r}

noaa_ports_complete %>% 
  filter(tresh == "75th") %>% 
  ggplot() +
  geom_point(
    aes(
      x = lon,
      y = lat,
      shape = species_name, 
      color = species_name
    ),
    size = 5,
    alpha = 0.5
  )

```

<!-- This is how all the shapefiles look like together -->

```{r all_sf, eval = T, echo = T}

# Visual exploration of all together (o.k! )
ggplot() +
  geom_sf(data = eez_sf, aes()) +
  geom_sf(data = land_sf, aes(fill = state), alpha = 0.3) +
  geom_sf(data = state_sf, aes(color = state), fill = NA) +
  geom_point(data = subset(noaa_ports_complete, tresh == "75th"), aes(x = lon, y = lat,
                                  shape = species_name), size = 3, alpha = 0.7) +
  scale_color_manual(values = state_pallet) +
  scale_fill_manual(values = state_pallet) +
  theme_dark() +
  coord_sf(ylim = c(33,48))

```

### Create base shapefile for computation

As a first step we need to divide the NE US EEZ among the different states. For that, we expanded a buffer-polygon to the 200 nautical miles to then estimate the percentage that each expanded polygon occupied. For the initial buffer point we used two approaches; state waters and US ports. Note that in all cases these areas overlapped and percentages accounted for it. We did this by following these steps:

- 1. Expand state waters / US ports using a buffer

- 2. Grid that buffer on a 0.3 resolution

- 3. Crop the buffer to the EEZ

#### 1. Expand Spatial regions (a.k.a  buffers)

##### 1.1 State Waters

We set a buffer of *410000* m (410 km, ~ 221 nm) that overshoots the EEZ a bit, but is eventually cropped

```{r sw_buffer, eval = T, echo = T}

# Buffer state waters
state_bf = st_buffer(state_sf, 4) %>% 
  st_transform(4326) %>% # to match shape
  st_set_crs(4326)

# ------------------------------ #
# Testing and visualizing the buffer 
# ------------------------------ #

state_bf_plot <- ggplot() +
  geom_sf(data = state_sf, aes(color = state), fill = NA) +
  geom_sf(data = eez_sf, aes(), fill = NA) +
  geom_sf(data = state_bf, aes(color = state), fill = NA) +
  scale_color_manual(values = state_pallet)+
  theme_dark()
#

# ggsave("../Results/Partial/state_waters_buffer.png",state_bf_plot)

# ------------------------------ #

state_bf_plot

```

##### 1.1 US. Ports

We set a buffer of *5* deg (510 km, ~ 221 nm) that overshoots the EEZ a bit, but is eventually cropped

```{r fp_buffer, eval = T, echo = T}

#Load top port data 
ports_df <- my_path("D","Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "75th")

port_sf <- st_as_sf(ports_df,
                    coords =c("lon","lat"),
                    crs = 4326)

# Buffer state waters
port_bf <- st_buffer(port_sf, 5) %>% # port
  st_transform(4326) %>% # to match shape
  st_set_crs(4326)

# ------------------------------ #
# Testing and visualizing the buffer 
# ------------------------------ #

ggplot() +
  geom_sf(data = eez_sf, aes(),fill = NA) +
  geom_sf(data = port_bf, aes(), fill = NA) +
  scale_color_manual(values = state_pallet)+
  theme_dark()

```

#### 2. Expand grid within buffer

Here we expand a grid within the buffer so we can estimate the proportion of each state.

*Note:* Dark grey shaded area is the grid and is the same for both approaches

```{r grid_indexing, eval = T, echo = T, message = F, warning = F}

# Create grid of the region
bbox <- c(st_bbox(state_bf))

# xmin      ymin      xmax      ymax 
# -82.57346  29.77459 -62.88543  49.19141

# Expand the grid
ne_grid <- expand.grid(
    lon = seq(from = bbox["xmin"], to = bbox["xmax"], by = 0.3),
    lat = seq(from = bbox["ymin"], to = bbox["ymax"], by = 0.3)
) %>% 
  rowid_to_column("index")

# ----------- #
# [TEST] Plot all layers
# ----------- #

# Looks good
state_sf %>%
  st_transform(4326) %>% # to match shape
  st_set_crs(4326) %>%
  # st_simplify(preserveTopology = TRUE, dTolerance = 10000) %>%
  ggplot() +
  geom_sf(aes(color = state))+
  # geom_sf(data = eez_sf, aes(),fill =NA) +
  # ggplot()+
  geom_tile(data = ne_grid,
            aes(
              x = lon,
              y = lat
            ),
            alpha = 0.2
  ) +
  scale_color_manual(values = state_pallet) +
  theme_bw() +
  theme_dark()

```

##### 2.1 Create grid for interpolation

```{r inter_grid, eval = T, echo = T}

# Crop grid to EEz
inter_grid_sf <- st_as_sf(ne_grid,
             coords = c("lon", "lat"),
             crs = 4326) %>% 
  st_intersection(eez_sf)

inter_grid_df <- as.data.frame(inter_grid_sf) %>% 
  select(index) %>% 
  left_join(ne_grid)

# Save grid for future analysis
# write_csv(inter_grid_df, paste0(my_path("R", "Partial"),"inter_grid_df.csv"))

ggplot(inter_grid_sf) +
  geom_sf()

```

##### 2.1 Merge grid and buffers

Once we have a gridded area, we converted the grid to a `sf` so we can merge it with the buffered states and ports and finally filter out everything outside the states polygon

###### 2.1.1 State Waters

```{r grid_sw_buf_merge, eval = T, echo = T, message = F, warning = F, fig.width = 10}

# ---------------- #
# Convert grid to sf
# ---------------- #
grid_sw_sf <- st_as_sf(ne_grid,
             coords = c("lon", "lat"),
             crs = 4326) %>% 
  st_join(state_bf) %>% 
  filter(!is.na(state))

# Create data frame for future computations
# Note, will be used in next chunk
grid_sw_bf_dt <- as.data.frame(grid_sw_sf) %>%
    select(index,state)
    # group_by(state) %>% 
    # summarise(length(index))

# ---------------- #
# [TEST] 
# Visualization of grid
# ---------------- #

gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_sw_sf) +
    geom_sf(data = state_sf, aes(), fill = NA) +
    geom_sf(aes(color = state), alpha = 0.3) +
    geom_sf(data = eez_sf, aes(),fill =NA) + 
    scale_color_manual(values = state_pallet) +
    theme_dark() +
    theme(legend.position = ""),
  # Showing each state separatley
  ggplot() +
    geom_sf(data = grid_sw_sf, aes(color = state),size = 0.1, alpha = 0.5) +
    facet_wrap(~state) +
    theme_dark() +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet),
  nrow = 1)

```

###### 2.1.1 Fishing Ports

```{r grid_fp_buf_merge, eval = T, echo = T, message = F, warning = F, fig.width = 10}

# ---------------- #
# Convert grid to sf 
# ---------------- #
grid_fp_sf <- st_as_sf(ne_grid,
             coords = c("lon", "lat"),
             crs = 4326) %>% 
  st_join(port_bf) %>% # join to port sf
  filter(!is.na(port_name)) #%>%
  # distinct(index,port_name, .keep_all = TRUE) #removes like 400K overlapping rows

# Create data frame for future computations
# Note, will be used in next chunk
grid_fp_bf_dt <- as.data.frame(grid_fp_sf) %>%
    select(index,port_name,species_name)

# ---------------- #
# [TEST] 
# Visualization of grid
# ---------------- #

gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_fp_sf) +
    geom_sf(data = port_sf, aes(), fill = NA) +
    geom_sf(aes(color = state_postal), alpha = 0.3) +
    geom_sf(data = eez_sf, aes(),fill =NA) + 
    scale_color_manual(values = state_pallet) +
    theme_dark() +
    theme(legend.position = ""),
  # Showing each state separately
  ggplot() +
    geom_sf(data = grid_fp_sf, aes(color = state_postal),size = 0.1, alpha = 0.5) +
    facet_wrap(~state_postal+species_name) +
    theme_dark() +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet),
  nrow = 1)

```

#### 3. Crop buffers to EEZ

Finally, we crop the grided buffers to within the EEZ to capture the actual water space.

*Note:* This step takes quite a while because of the size of the EEZ shapefile. No, you cannot use `st_simplify()` here

#### 3.1 State Waters

```{r buff_sw_to_eez, eval =T, echo = T, message = F, warning = F}

# I don't know how to undo st_simplify so need to re-load the shapefile

eez_sf <- st_read(my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile", name = "SAUEEZ_July2015.shp")) %>% 
  rename(eez_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  filter(eez_name == "USA (East Coast)") %>%

# Get the overlapping segments 
# NOTE: Takes some gooooood time (~20 min)
Sys.time()
grid_eez_sw_sf <- st_intersection(grid_sw_sf,eez_sf)
Sys.time()

# write_sf(grid_eez_sw_sf, paste0(my_path("D", "Spatial/grid_eez_sw_sf"),"grid_eez_sw_sf.shp"))
grid_eez_sw_sf <- my_path("D","Spatial/grid_eez_sw_sf", name = "grid_eez_sw_df.csv",read = T)

# Get final df with index
grid_eez_sw_df <- as.data.frame(grid_eez_sw_sf) %>%
  select(state,abrev,index) %>%
  left_join(ne_grid,
            by = "index")

# write_csv(grid_eez_sw_df, paste0(my_path("D", "Spatial/grid_eez_sw_sf","grid_eez_sw_df.csv")))

# Plot to make sure makes sense

eez_sf <- eez_sf %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) #0.1 for paper

grid_eez_sw_df %>%
  ggplot() +
  geom_sf(aes(color = state), alpha = 0.3) +
  scale_color_manual(values = state_pallet) +
  geom_sf(data = eez_sf,aes(), fill = NA) +
  theme_dark()

```

#### 3.2 Fishing Ports

```{r buff_fp_to_eez, eval =T, echo = T, message = F, warning = F}

state_names <- grid_eez_sw_df %>% 
  select(state,abrev) %>% 
  distinct(.keep_all = T)

# Get the overlapping segments 
# NOTE: Takes some gooooood time (~ min)
Sys.time()
grid_eez_sf_fp <- st_intersection(grid_fp_sf,eez_sf)
Sys.time()

grid_eez_sf_fp_save <- grid_eez_sf_fp %>% 
  select(spp = species_name,
         index,
         port_name,
         abrev = state_postal,
         geometry) %>% 
  left_join(as.data.frame(land_sf),
            by = "abrev") %>% 
  select(-geom,-ID)

write_sf(grid_eez_sf_fp_save, my_path("D", "Partial/grid_fp",name="grid_eez_fp.shp"))

# Get final df with index
grid_eez_fp_df <- as.data.frame(st_coordinates(grid_eez_sf_fp_save)) %>% 
  rename(lon = X, lat = Y) %>% 
  bind_cols(grid_eez_sf_fp_save) %>% 
  select(-geometry)

write_csv(grid_eez_fp_df, my_path("D", "Partial/grid_fp",name = "grid_eez_fp_df.csv"))

# Plot to make sure makes sense
grid_eez_sf_fp %>%
  ggplot() +
  geom_sf(aes(color = state_postal), alpha = 0.3) +
  scale_color_manual(values = state_pallet) +
  geom_sf(data = eez_sf,aes(), fill = NA) +
  theme_dark()

```

#### 3.3 Differences at croped level

By looking at the plots we do not see that the number of grid cells that each state has is slightly different depending on the approach. This comes to light when we count the number of gridcells that each state gets in each approach. We see that the state waters one is substantially higher.

We can visualize the difference of the cropped area for each state next. Again, the fishing port approach results in less area for all states, but such area is not proportional among states. For example, Delaware looses roughly 35% of fishing grounds with this approach

```{r state_bf_diff_ntbl, eval = T, echo = T}

my_path("R", "Partial",name = "grid_eez_sw_sf.csv", read = T) %>% 
  mutate(approach = "state_waters") %>% 
  bind_rows(my_path("R", "Partial/grid_fp",name = "grid_eez_fp_df.csv", read = T)) %>%
  mutate(approach = ifelse(is.na(approach),"fishing_port",approach)) %>% 
  group_by(state,approach) %>% 
  summarise(n_grid = n()) %>% 
  spread(approach,n_grid) %>% 
  mutate(difference =state_waters-fishing_port,
         percentage_sw = round((difference/state_waters)*100)
         )

```

And here is the map visualization of the cropped difference

```{r state_bf_diff_map, eval = T, echo = T}

grid_eez_sw_df %>% 
  mutate(approach = "state_waters") %>% 
  bind_rows(grid_eez_fp_df) %>%
  mutate(approach = ifelse(is.na(approach),"fishing_port",approach)) %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = approach
    ),
    alpha = 0.5
  ) +
  facet_wrap(~state) +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2") +
  theme_dark()

```

### Interpolation rutine

In this step we [interpolate](https://swilke-geoscience.net/post/spatial_interpolation/) the survey data within the previously created grid following a Triangular Irregular Surface method.

#### Functions needed

We need to create a couple of functions to run the whole process

##### Interpolation main fx (tis).

This is the main function used to interpolate the data per year. It follows a Triangular Irregular Surface method using the `interp::interp()` function. If you want to see the function clic on `code`

```{r interpol_function, eval = T, echo = T}

tis <- function(input_data, grid, yr, taxa, reg){

  # --------------- #
  # Testing
  # print(paste(yr))
  # yr = 1976
  # --------------- #

  # Filter data
  data <- input_data %>% 
    filter(year == yr,
           spp == taxa,
           region == reg
    ) %>% 
    # Average duplicated hauls in the same spot
    group_by(region,year,spp,lat,lon) %>%
    summarise_at(vars(wtcpue),
                 mean,na.rm = T)

  # Only interpolate cases where there is more than 3 rows
  # Marked by the function 
    if(nrow(data) <= 3){
      fit_tin <- tibble()
    }else{

      # Triangular Irregular Surface
      fit_tin <- interp::interp( # using {interp}
        x = data$lon,           # the function actually accepts coordinate vectors
        y = data$lat,
        z = data$wtcpue,
        xo = grid$lon,     # here we already define the target grid
        yo = grid$lat,
        output = "points"
      ) %>% 
        bind_cols() %>% 
        bind_cols(grid) %>%
        mutate(year = yr,
               region = reg,
               spp = taxa) %>% 
        select(index, 
               # state,
               year, region, spp, lon=x, lat=y, value = z)

    }

    return(fit_tin)
}

# Test me
# Needs variables in Control panel
# Test no data: "Alosa aestivalis", reg = "Northeast US Fall", yr = 1974 
# tis(input_data = ocean_data, grid = grid_eez_df, taxa = "Illex illecebrosus", reg = "Northeast US Fall", yr = 1973)

# lapply(years,tis,input_data = ocean_data, grid = grid_eez_df, taxa = "Illex illecebrosus", reg = regions[2])

```

##### Run function

This is a sub-function that runs the `tis()` function by taxa and region. It saves the output as a .csv file for each species.

```{r interpol_run_fun, eval = T, echo = T}

run_tis <- function(input_data, grid, years, taxa, reg){

  # Run tis for species and surveys
  for(r in 1:2){

    partial_df <- bind_rows(
      lapply(years,tis,input_data = input_data, grid = grid, taxa = taxa, reg = reg[r])
    )

    if(r == 1){
      historic_tif <- partial_df
    }else{
      historic_tif <- bind_rows(historic_tif,partial_df)
    }

  }

  # ----------------------- #
  # Save dataset per species
  # ----------------------- #

  # Set file name
  name <- paste0("tif_",str_replace(taxa," ","_"),".csv")

  complement <- paste0("Partial/Interpolation/")

  # Set path name
  save_path <- my_path("R",complement)

  # Set complete path
  save_name <- paste0(save_path,name)

  # Create folder if it does not exist
  if(file.exists(save_path) == F){
    dir.create(save_path)
  }

  #  Save file
  write_csv(historic_tif,save_name)

  return_msg <- paste("Interpolation done for", taxa)

  return(return_msg)

}

# Test me
    # run_tis(input_data = ocean_data, grid = grid_eez_fp_df, taxa = "Centropristis striata", years = seq(1970,1980), reg = regions, method = "testa")

```

#### Survey data

The interpolation was done with NOAA Northeast Fisheries Science Center Spring and Fall Bottom Trawl Surveys [data](https://www.fisheries.noaa.gov/region/new-england-mid-atlantic#science) provided by [Ocean adapt](https://oceanadapt.rutgers.edu/). Data was accessed trough the [Github](https://github.com/pinskylab/OceanAdapt).

*In primary publications using data from the database, please cite Pinsky et al. 2013. Marine taxa track local climate velocities. Science 341: 1239-1242 doi: 10.1126/science.1239352, as well as the original data sources.*

- Using only the Northeast US Fall and Spring bottom trawl survey data for now

##### Splitting up data

- No part on spatial analysis. Can be ignored.

This is just a sub-step to split up the data into single species files. This makes the app faster as it only needs to load species specific data, rather than all the data at de beginning.

```{r dat_exploded, eval = F, echo = F, fig.width = 9}

ocean_data <- readRDS("/Volumes/Enterprise/Data/pinskylab-OceanAdapt-966adf0/data_clean/dat_exploded.rds") #%>% 
  # filter(spp == "Centropristis striata",
       # region %in% c("Northeast US Fall" , "Northeast US Spring")) #No more seasons

spp_data <- function(spp){

  name_save <- paste0(my_path("D","Spp/Observation"),"obs_",str_replace(spp[1], " ", "_"),".csv")

ocean_data %>% 
  filter(spp == spp,
       region %in% c("Northeast US Fall" , "Northeast US Spring")
       ) %>% 
  write_csv(., name_save)

}

spp_list <- ocean_data %>% 
  filter(region %in% regions,
         spp != "NA") %>% 
  pull(spp) %>% 
  unique()

lapply(spp_list, spp_data)

subset_data <- ocean_data %>% 
  filter(spp == "Centropristis striata",
       region %in% c("Northeast US Fall" , "Northeast US Spring")
       )

ggplot() +
  geom_point(data = subset(subset_data, wtcpue = 0.0),
             aes(
               x = lon,
               y = lat
             ),
             color = "grey95"
  ) +
  geom_point(data = subset(subset_data, wtcpue > 0),
             aes(
               x = lon,
               y = lat,
               color = log10(wtcpue)
             ),
             size = 1
  ) +
  scale_color_distiller(palette = "Spectral", 
                        guide_legend(title = "WCPUE per Haul (log10)")) + 
  coord_sf(xlim = c(-76, -65),ylim = c(35, 45)) +
  MyFunctions::my_ggtheme_m() +
  facet_wrap(~region)

```

#### Control Pannel

This is where we load the required data and prepare to run the interpolation function. Note that some of the data here has been previously created

```{r contro_pannel, eval = T, echo = T}

# Load grid df
# For fishing ports
inter_grid <- my_path("D","Spatial","inter_grid_df.csv", read = T)

# Run interpolation for all years
years <- seq(1973,2019,1)

# regions
regions <- c("Northeast US Fall" , "Northeast US Spring")

# species list
spp <- ocean_data %>% 
  filter(region %in% regions,
         spp != "NA") %>% 
  pull(spp) %>% 
  unique()

# Double check runs

spp_runs <- tibble(taxa = (list.files(my_path("R","Partial/Interpolation/fp")))) %>%
  mutate(
    taxa = str_remove(taxa, "tif_"),
    taxa = str_remove(taxa, ".csv"),
    taxa = str_replace(taxa, "_", " ")
  )

# Missing runs
spp <- tibble(taxa=spp) %>% 
  anti_join(spp_runs) %>% 
  pull(taxa)

```

#### Run routine

Here we just run the routine for each of the species present in the Northeast US Fall and Spring surveys between 1973 and 2019.

- Note there are 43 identified taxa
- Some taxa do not have presence data in some years

```{r run_routine, eval = T, echo = T}

# # single species run
run_tis(input_data = ocean_data,
        grid = grid_eez_fp_df,
        years = years,
        reg = regions,
        taxa = "Stenotomus chrysops",
        method = "fp"
)

# Run them all in parallel
lapply(Species,
       run_tis, 
       input_data = ocean_data,
       grid = inter_grid,
       years = years,
       reg = regions
)

```

# Results

Results are now for eight species managed under Mid-Atlantic Council Management Plans according to [NOAA](https://www.fisheries.noaa.gov/new-england-mid-atlantic/commercial-fishing/new-england-mid-atlantic-fishery-management-plans).

State allocations of the commercial black sea bass coastwide quota were originally implemented in 2003 as part of Amendment 13, loosely based on historical landings from 1980- 2001.

*Scomber scombrus*, *Peprilus triacanthus* (butterfish), *Illex illecebrosus* (shortfin squid), *Paralichthys dentatus* (summer flounder), *Stenotomus chrysops* (Scoop), *Centropristis striata* (black sea bass), *Pomatomus saltatrix* (Bluefish), *Lopholatilus chamaeleonticeps* (Golden tilefish), *Caulolatilus microps* (blueline tilefish) and *Clupea harengus*

```{r}

x <- as.data.frame(st_read(my_path("D", "Partial/grid_fp",name="grid_eez_fp.shp"))) %>% 
  select(index,spp = spcs_nm,port_name = port_nm, abrev = stt_pst) %>% 
  bind_cols(
    as.data.frame(st_coordinates(st_read(my_path("D", "Partial/grid_fp",name="grid_eez_fp.shp"))))
  ) %>% 
  rename(lon = X, lat = Y) %>% 
  left_join(as.data.frame(land_sf),
            by = "abrev") %>% 
  mutate(state= str_to_sentence(ID)) %>% 
  select(1:6,state) %>% 
  mutate(
     spp = ifelse(spp == "summer flounder","Paralichthys dentatus", 
                 ifelse(spp == "scup", "Stenotomus chrysops", 
                        "Centropristis striata"
                 )
    )
  )

write_csv(x,my_path("D", "Partial/grid_fp",name="grid_eez_fp_df.csv"))

# x %>% filter() %>% group_by(spp,abrev) %>% tally()

```

```{r load_data, eval = T, echo = T}

# Spatial data
States <- c("maine", "new hampshire", "massachusetts", "connecticut", "rhode island", "new york", "new jersey", "delaware", "maryland", "virginia", "north carolina")

# US State Map (land)
land_sf <- st_as_sf(map("state", plot = FALSE, fill = TRUE)) %>% 
  filter(ID %in% States) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA"),
         state = str_to_sentence(ID)
         )

# US EEZ shapefile
eez_sf <- st_read(my_path("G", extra_path = "Spatial/SAU/SAU_Shapefile", name = "SAUEEZ_July2015.shp")) %>% 
  st_transform(crs = 4326) %>% # 4326
  filter(Name == "USA (East Coast)") %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) #0.1 for paper

### Previously created grids

grids_sw <- bind_rows(
  my_path("D", "Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T) %>% 
    mutate(spp = "Paralichthys dentatus"),
  my_path("D", "Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T) %>% 
    mutate(spp = "Stenotomus chrysops"),
  my_path("D", "Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T) %>% 
    mutate(spp = "Centropristis striata")
) %>% 
  left_join(land_sf) %>% 
  select(index,state,abrev,lat,lon,spp) %>% 
  mutate(spatial = "sw",
         port_name = NA
         )

grids_fp <- my_path("D", "Partial/grid_fp", name = "grid_eez_fp_df.csv", read = T
) %>%
  mutate(spatial = "fp",
  spp = ifelse(spp == "summer flounder","Paralichthys dentatus", 
                 ifelse(spp == "scup", "Stenotomus chrysops", 
                        "Centropristis striata"
                 )
    )
  )

# Join both approaches
grids <- bind_rows(grids_sw,grids_fp)

# unique(grids$spp)

# Ports data
ports_df <- my_path("D","Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "75th") %>%
  group_by(species_name,state_postal) %>% 
  summarise(n =n()) %>% 
  select(-n)
  # distinct(c(,state_postal), .keep_all = T) # No need for duplication

# Species data

# Managed species by Mid-Atlantic Council Management Plans, and State waters
# https://www.fisheries.noaa.gov/new-england-mid-atlantic/commercial-fishing/new-england-mid-atlantic-fishery-management-plans

Species <- c(
  "Paralichthys dentatus", #summer flounder
  "Stenotomus chrysops", #Scup"
  "Centropristis striata"#, # black sea bass
  )

spp_paths <- list.files(my_path("R",extra_path = "Partial/Interpolation/"))

# Some gibiris for names
taxon_list <- gsub("_"," ",spp_paths)
taxon_list <- gsub("tif ","",taxon_list)
taxon_list <- gsub(".csv","",taxon_list)

taxon_list <- taxon_list[taxon_list %in% Species]

# Now set a list of files to read
taxon_read <- paste0("tif_",taxon_list,".csv")
taxon_read <- gsub(" ","_",taxon_read)

taxon_read <- spp_paths[spp_paths %in% taxon_read]

#### Load sw interpolation

taxon_sw <- my_path("R",extra_path = "Partial/Interpolation/", name = taxon_read)

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat","spp")
  ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall"),
    label = ifelse(spp == "Paralichthys dentatus" & year >= 1980 & year <= 1986,"Reference",
                   ifelse(spp == "Stenotomus chrysops" & year >= 1988 & year <= 1992,"Reference",
                          ifelse( spp == "Centropristis striata" & year >= 1980 & year <= 2001,"Reference",
                                 ifelse(year > 2001,"Today",NA)
                          )
                   )
    )
  )

# historic_spp %>% 
#   filter(is.na(spatial)) %>% 
#   filter(year == 1973,
#          region == "Northeast US Fall") %>% 
#   ggplot() +
#     geom_tile(
#       aes(
#         x = lon,
#         y = lat,
#         fill = index
#       )
#     ) +
#   facet_grid(spp~state) + 
#   geom_sf(data = land_sf,aes()) +
#   geom_sf(data = eez_sf, aes(), fill =NA)

# Data summary
c(unique(historic_spp$spp),
  unique(historic_spp$region),
  unique(historic_spp$season),
  unique(historic_spp$spatial)
)

# Periods
periods <-tibble(
  order = c(rep("a",12),
            rep("b",12),
            rep("c",12),
            rep("d",11)
  ),
  label = c(rep("1973-1984",12),
            rep("1985-1996",12),
            rep("1997-2008",12),
            rep("2009-2019",11)
  ),
  year = c(seq(1973,1984,1),
           seq(1985,1996,1),
           seq(1997,2008,1),
           seq(2009,2019,1)
  )
  )

state_lat <- historic_spp %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  filter(!is.na(state)) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA"),
         order = ifelse(abrev == "ME", order+1,order)
  ) %>% 
  arrange(desc(order))

# State pallet

state_pallet <- c(wes_palette(n = 5, name = "Darjeeling1"),
                               wes_palette(n = 5, name = "Darjeeling2"),
                               # wes_palette(n = 2, name = "GrandBudapest1")
                  "#FD6467"
                  )

# print for notebook
head(historic_spp)
```

## Regulatory units result

```{r cropped_plot_paper, eval = F, echo = T}

#### Data needed ####

# Load grid of state waters
grid_eez_sw_sf <-  st_read(my_path("D","Partial/grid_sw", name = "grid_eez_sw_sf.shp")) %>%
  select(index,state,geometry) %>% 
  mutate(method = "state_waters",
         state = str_to_sentence(state)) %>% 
  left_join(state_lat)

grid_eez_sw_sf$group <- factor(grid_eez_sw_sf$abrev,      # Reordering group factor levels
                               levels = state_lat$abrev)

# Load grid of fishing ports
grid_eez_fp_sf <-  st_read(my_path("D","Partial/grid_fp", name = "grid_eez_fp.shp")) %>%
  select(index,abrev,geometry) %>% 
  mutate(method = "fishing_port"#,
         # state = str_to_sentence(landing_port)
         )%>% 
  left_join(state_lat)

grid_eez_fp_sf$group <- factor(grid_eez_fp_sf$abrev,      # Reordering group factor levels
                               levels = state_lat$abrev)

grid_eez_fp_sf <- arrange(grid_eez_fp_sf,group)
# Plotting 
# Make sure st_simplify() is run for eez sf

# State waters plot

p_sw <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color =group , fill = group), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45))+
    labs(x = "", y = "", title = "A) State waters approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf(aes(color = group),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_sw_sf$group)) +
                       labels = grid_eez_sw_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

##### ------------- ####
# Fishing Ports plot
##### ------------- ####

grid_eez_fp_sf <- grid_eez_sw_sf %>% 
  filter(!group %in% grid_eez_fp_sf$group) %>% 
  mutate(index = NA) %>% 
  bind_rows(grid_eez_fp_sf) %>% 
  distinct(.keep_all = T)

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(data = subset(grid_eez_fp_sf, index != "NA")) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = group, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    # scale_color_manual(values = state_pallet) +
    # scale_fill_manual(values = state_pallet) +
    scale_color_manual(values = c(state_pallet[3:7],state_pallet[9:11])) +
    scale_fill_manual(values = c(state_pallet[3:7],state_pallet[9:11])) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    labs(x = "", y = "", title = "B) Fishing ports approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot() +
    geom_sf(data = subset(grid_eez_fp_sf, index = "NA"), color = "white") +
    geom_sf(data = subset(grid_eez_fp_sf, index != "NA"), aes(color = group),size = 0.1, alpha = 0.3) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    facet_wrap(~ group) +
    # geom_point(data = rename(ports_df, group = state_postal), aes(x = lon, y = lat), color = "#E6A0C4") +
    theme(legend.position = "top") +
    # scale_color_manual(values = state_pallet,
    scale_color_manual(values = c(state_pallet[3:7],state_pallet[9:11]),
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

combined_plot <- gridExtra::grid.arrange(p_sw,p_fp,
                                         bottom = "Longitude",
                                         left = "Latitude")

ggsave(filename = "buffer_figure.jpg",
       plot = combined_plot,
       path = my_path("R","Figures"),
       width = 11,
       height = 11
       )

```

## Interpolation result

```{r}

# Interpolation data
interpol_result <- historic_spp %>% 
  filter(!is.na(label)) %>% 
  group_by(label,index,spp,season,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue))

# Paper numbers

# interpol_result %>% 
#   filter(mean_cpue >0) %>% 
#   View()

# --------------------- #
# Biomass Centroid shift
# --------------------- #

# Estimate total interpolation
total_inter <- historic_spp %>% 
  filter(!is.na(label)) %>% 
  group_by(spp,index,season) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  group_by(spp,season) %>% 
  summarise(total_cpue = sum(mean_cpue, na.rm = T))

# Estimate centroid based on top 10 grids with higer abundance
centroid_db <- historic_spp %>% 
  filter(!is.na(label)) %>% 
  group_by(label,index,spp,season,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue)) %>% 
  left_join(total_inter) %>% 
  mutate(
    per_cpue = round((mean_cpue/total_cpue)*100,2)
  ) %>% 
  group_by(spp,label,season) %>% 
  top_n(10,per_cpue) %>% 
  group_by(spp,label,season) %>% 
  summarise(
    lat = mean(lat),
    lon = mean(lon)
  )

# Paper numbers
# centroid_db %>% 
#   select(-lon) %>% 
#   spread(label,lat) %>% 
#   mutate(diff = Today-Reference) %>% 
#   View()

# For alternative color pallet
pal <- wes_palette("Zissou1",100,type = "continuous")

inter_map <- ggplot() +
  geom_tile(data = subset(interpol_result, mean_cpue == 0),
            aes(
              x = lon,
              y = lat
            ),
            fill = "grey90",
            color = "grey90"
  ) + 
  geom_tile(data = subset(interpol_result, mean_cpue > 0),
            aes(
              x = lon,
              y = lat,
              fill = log10(mean_cpue),
              color = log10(mean_cpue)
              # fill = mean_cpue,
              # color = mean_cpue
            )
  ) + 
  geom_point(data = centroid_db,
             aes(
               x = lon,
               y = lat
             ),
             shape = 3,
             color = "white"
  ) + 
  geom_sf(data = land_sf,aes()) +
  labs(x = "Longitude",
       y = "Latitude") +
  # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
  #                                  size = 4,
  #                                  box.padding = 0.10,
  #                                  hjust = 1) +
  # Viridis option
# scale_fill_viridis("CPUE",
#                    limits = c(0,40),
#                    # limits = c(-5,1), # for log10
#                    direction = -1,
#                     end = 0.9
#                    ) +
# scale_color_viridis("CPUE",
#                     limits = c(0,40),
#                     # limits = c(-5,1), # for log10
#                     direction = -1,
#                     end = 0.9
#                     ) +
# wes anderson option
scale_fill_gradientn("CPUE",
                     colours = pal,
                     # limits = c(0,40) # For cpue
                     limits = c(-5,1) # for log10

) +
  scale_color_gradientn("CPUE",
                        colours = pal,
                        # limits = c(0,40) # for cpue
                        limits = c(-5,1) # for log10
  ) +
  facet_grid(spp~season+label) +
  my_ggtheme_m(map_type = "na", leg_pos = "right",ax_tx_s = 7,ax_tl_s = 10,leg_tx_s = 12) #+
# theme(legend.key.width = unit(3,"line"))

ggsave("interpolation_wa.jpg",
       inter_map,
       path = my_path("R","Figures"),
       width = 9,
       height = 7)

```

## Average proportion

This map shows the aggregated extrapolated value from all three species per State average across the whole study period within each State's water.

*Note:* This is intended to be a supplemental figure

```{r area_map_state, eval = T, echo = T, fig.width=10, fig.height=12}

data_grid <- historic_spp %>% 
  group_by(state,lat,lon,region,spatial) %>% 
  summarise(sum = sum(value,na.rm= T), .groups = "drop") %>% # sum of all species
  group_by(state,lat,lon,region,spatial) %>% 
  summarise(mean = mean(sum,na.rm= T), .groups = "drop") # average of all years

ggplot(data = subset(data_grid, !is.na(mean))) +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = mean,
      col = mean
    )
  ) +
  geom_sf(data = land_sf, aes()) +
  facet_wrap(~ state  +  region + spatial) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_viridis("Mean Interpolation", alpha = 0.8) +
  scale_color_viridis("Mean Interpolation", alpha = 0.8) +
  theme_bw()
```

## Proportion Change

Here we show the average proportion of the interpolation by State and time period. Time periods were arbitrary defined as;

- Early; 1973 to 1984
- Mid; 1985 to 1997
- Late; 1998 to 2011
- Now; 2012 to 2019

*Notes:* Figure represents the **Spring** survey. This computation considers the Overlapping of state waters.

### Data preparation

```{r data_prep, eval = F, echo = T, results='hide'}

total_fited <- historic_spp %>% 
  group_by(year,season,spp,spatial) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_fit <- historic_spp %>% 
  group_by(state,year,label,season,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  # View()
  left_join(total_fited,
            by = c("year","season","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,label,season,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(!is.na(label))

```

#### F3 - Regulatory units map

```{r reg_unit_map_diff, eval = F, echo = T, results='hide'}

reg_unit <- state_fit %>% 
  filter(season == "Spring") %>% 
  spread(spatial,mean_per) %>% 
  # Set NA's to 0 because no data found on FP
  mutate(fp = replace_na(fp,0)) %>%
  # View()
  mutate(
    difference = fp-sw,
    difference = ifelse(difference > 100,100,
                        ifelse(difference < -100,-100,difference)
    )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(spp = gsub(" ","\n",spp),
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State waters","Difference")
         )
  )

# The plot
map_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,35,5),
                              limits=(c(0,35))
                              )+
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(reg_unit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  # viridis::scale_fill_viridis("Difference (%)", alpha = 0.8,option = "C",
  #                             breaks = seq(0,25,5),
  #                             limits=(c(0,25)),
  #                             # labels = c("<-10",seq(-75,75,25),">100")
  #                             ) +
  scale_fill_gradient2("% Difference\n(fp-sw)", 
                       breaks = seq(-14,35,7),
                       limits=(c(-14,35)),
                              # labels = c("<-10",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

# Cowplot option
p <- ggdraw() +
  # Revenue circular
  draw_plot(map_plot, x = 0, y = 0.2, width = 1, height = 0.8) +
  # Catch circular
  draw_plot(diff_plot, x = 0, y = 0, width = 1, height = 0.4) +
  draw_plot_label(label = c("Latitude", "Longitude"), 
                  size = 18,
                  angle = c(90,0),
                  x = c(0,0.45), 
                  y = c(0.45,0.15)
  )

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_reg.jpg",
  p,
  base_height = 14,
  base_width = 14
)

```

#### F4 - Seasonal units map

```{r sea_unit_map_diff, eval = F, echo = T, results='hide'}

# Get seasonal data
season_fit <- state_fit %>% 
  filter(spatial != "fp") %>%
  spread(season,mean_per) %>%
  mutate(
    Difference = Fall-Spring,
    Difference = ifelse(Difference > 100,100,
                        ifelse(Difference < -100,-100,Difference)
    )
  ) %>% 
  gather("season","mean_per",Fall:Difference) %>% 
  mutate(spp = gsub(" ","\n",spp))

# The plot
map_plot <- land_sf %>% 
  left_join(season_fit,
            by = "state") %>% 
  filter(season != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,15,5),
                              limits=(c(0,15))
                              )+
  facet_grid(season~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(season_fit,
            by = "state") %>% 
  filter(season == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  # viridis::scale_fill_viridis("Difference (%)", alpha = 0.8,option = "C",
  #                             breaks = seq(0,5,1),
  #                             limits=(c(0,5)),
  #                             # labels = c("<-10",seq(-75,75,25),">100")
  #                             ) +
  scale_fill_gradient2("% Difference\n(Fall-Spring)", 
                       breaks = seq(-4,4,2),
                       limits=(c(-4,4)),
                              # labels = c("<-10",seq(-75,75,25),">100")
                              ) +
  facet_grid(season~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

# Cowplot option
p <- ggdraw() +
  # Revenue circular
  draw_plot(map_plot, x = 0, y = 0.2, width = 1, height = 0.8) +
  # Catch circular
  draw_plot(diff_plot, x = 0, y = 0, width = 1, height = 0.4) +
  draw_plot_label(label = c("Latitude", "Longitude"), 
                  size = 18,
                  angle = c(90,0),
                  x = c(0,0.45), 
                  y = c(0.45,0.15)
  )

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_seas_.jpg",
  p,
  base_height = 14,
  base_width = 14
)

```

## F5 - Area plot

This graph shows the proportion of the interpolation value each State has over time.

### Area plot (running mean)

This graph shows the proportion of each State smoothed over a *10 years running mean**. It allows to better see increasing and decreasing trends.

#### Data preparation

```{r area_plot_Data_prep}

total_fited <- historic_spp %>% 
  group_by(year,region,spp,spatial) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  # View()
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spp,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State waters","Difference")
         )
  )

```

### Paper figure: Per Species Proportional change

```{r paper_plot_area_ave_spp, eval = F, echo = F, results='hide'}

# Shaded period 
historic_periods <- historic_spp %>% 
  filter(label == "Reference") %>% 
  group_by(spp,label,year) %>% 
  summarise(n = n()-n()+100) %>% 
  mutate(period = "Reference period")

# Plot me
ssp_area <-
  ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  geom_area(data = historic_periods,
             aes(
               x = year,
               y = n,
               color = period
             ),
             # fill = "gray90",
             alpha = 0.3,
            stat = "identity"
             ) +

  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State\n(Latitudinal order)",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
    scale_color_manual("",values = "grey90"
  ) +
  facet_grid(spp~spatial) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 0.9,ax_tx_s = 14,leg_tx_s = 16, leg_tl_s = 18,ax_tl_s = 18,facet_tx_s = 16) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5)

  ggsave(filename = "area_plot_avg_spp_sptl.jpg",
         plot = ssp_area,
         path = my_path("R","Figures"),
         width = 14,
         height = 8
  )

```
<!-- ![Proportion change aggregated all species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/area_plot_avg_spp_sptl.jpg) -->

# Old code

## Fishing ports WPI

We used [Global Fishing Watch](https://globalfishingwatch.org/datasets-and-code-anchorages/)'s database on 
Anchorages, Ports and Voyages.

*The Global Fishing Watch anchorages dataset is a global database of anchorage locations where vessels congregate. The dataset contains over 160,000 individual anchorage locations, which are further grouped into nearly 32,000 ports when applicable.*

```{r ports_sf, eval = T, echo = T}

# Get Global Fishig Watch data
gfw_data <- my_path("D","Spatial",name = "gfw_portdata.csv", read = T) %>% 
  filter(iso3 == "USA")

# Visualize data (o.k!)
# ggplot(gfw_data) +
#   geom_point(
#     aes(
#     x = lon,
#     y = lat
#   )
#   )

# Convert data to sf for mergig
gfw_sf <- st_as_sf(gfw_data,
             coords = c("lon", "lat"),
             crs = 4326)

# Merge data with state waters to,locate ports 
port_sf <- st_join(gfw_sf,state_sf) %>% 
  filter(!is.na(jurisdicti)) # remove points outside stat waters

ggplot(port_sf)  +
  geom_sf(aes(color = distance_from_shore_m)) +
  theme_dark()

```
So...Looks like some ports are "far from shore" which I don't really know what it means. For now, I am only selecting those ports that ar 0 meter from shore.

```{r ports_inshore, eval = T, echo = F}

# Get number of ports per state
ports_n_data <- port_sf %>% 
  as.data.frame() %>% 
    filter(distance_from_shore_m == 0) %>% 
    group_by(state) %>% 
    summarise(n_ports = n())

# plot it
gridExtra::grid.arrange(
  port_sf %>% 
    filter(distance_from_shore_m == 0) %>% 
    ggplot()  +
    geom_sf(aes(color = jurisdicti)) +
    scale_color_manual(values = state_pallet) +
    theme_dark() +
    theme(legend.position = ""),
  tableGrob(ports_n_data),
  ncol = 2
)

```

There are still too many ports in each state (?). Specially New Jersey! One approach is to use NOAA's landing-per-port data to filter out...

###### 4.2 World Port Index

In this approach we only use the ports categorized within the [WPI](https://msi.nga.mil/NGAPortal/MSI.portal?_nfpb=true&_pageLabel=msi_portal_page_62&pubCode=0015) and with `distance_from_shore_m == 0`  and `at_dock == TRUE`.

- `at_dock` == Anchorages within a 450 meter buffer around a combined land shape file are considered at dock.

```{r wpi_approach, eval = T, echo = F}

gridExtra::grid.arrange(
  port_sf %>% 
    filter(distance_from_shore_m == 0,
           label_source %in% c("WPI_ports"),
           at_dock == TRUE) %>% 
    ggplot()  +
    geom_sf(aes(color = jurisdicti)) +
    scale_color_manual(values = state_pallet) +
    theme_dark(),
  port_sf %>% 
    as.data.frame() %>% 
    filter(distance_from_shore_m == 0,
           label_source %in% c("WPI_ports"),
           at_dock == TRUE) %>% 
    group_by(state) %>% 
    summarise(n_ports = n()) %>% 
    tableGrob(),
  ncol = 2
)

```

## Historic quota Vs distributions

This section of the results explores the differences between the historic distribution of te stock and the historic quota allocation. We collected the proportion of the total quota that each State gets for each species.

Right now the analysis only covers black sea bass, summer flounder and scup. But we can go speceis by species to see which ones have their quota allocated by state to include them in the analysis.

- [Black Sea Bass; *Centropristis striata*](https://www.mafmc.org/newsfeed/2021/asmfc-and-mafmc-approve-bsb-commercial-allocation-changes)

**Note:** We are using the new allocations based on the stock's most recent biomass distribution*

- [Summer flounder](http://www.asmfc.org/uploads/file/5f9af82a2019SummerFlounderFMP_Review.pdf)

- [Scup](http://www.asmfc.org/uploads/file/5f99d0112019_Scup_FMP_Review_approved.pdf) Summer period

For State-level managed species see the [Atlantic States Marine Fisheries Comission](http://www.asmfc.org/fisheries-management/program-overview) website and go species-by-species.

```{r state_quota_tbl,eval= T, echo = T}

quota_allocation <- state_lat %>% 
  mutate(
    "Centropristis striata" = c(
      0.0040, #ME
      0.0040, #NH
      0.1564, #MA
      0.1323, #RI
      0.0367, #CT
      0.0857, #NY
      0.2010, #NJ
      0.0411, #DE
      0.0888, #MD
      0.1614, #VA
      0.0888 #NC
    ),
    "Paralichthys dentatus" = c(
      0.0004756, #ME 
      0.0000046, #NH 
      0.0682046, #MA 
      0.1568298, #RI 
      0.0225708, #CT 
      0.0764699, #NY 
      0.1672499, #NJ 
      0.0001779, #DE 
      0.0203910, #MD
      0.2131676, #VA
      0.2744584 #NC
    ),
    "Stenotomus chrysops" = c(
      0.0012101, #ME
      0.0000000, #HN
      0.2158729, #MA
      0.5619456, #RI
      0.0315399, #CT
      0.1582466, #NY
      0.0291667, #NJ
      0.0000000, #DE
      0.0001190, #MD
      0.0016502, #VA
      0.0002490 #NC
    )
  ) %>% 
  gather("spp","quota",4:6)

# Double check they all add to 1...
quota_allocation %>% 
  group_by(spp) %>% 
  summarise_at(vars(quota),sum)

quota_allocation %>% 
  arrange(order) %>% 
  select(-order) %>% 
  spread(spp,quota) %>% 
  kable()
```

### Get data ready

```{r}

total_fited <- historic_spp %>% 
  group_by(year,region,spp) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spp) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"))

quota_df <- quota_allocation %>% 
  left_join(historic_spp) %>% 
  group_by(state,abrev,year,region,spp,quota) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp")) %>%
  mutate(Distribution = state_value/total_value*100,
         Historic = quota*100) %>% 
  group_by(state,quota,spp) %>% 
  mutate(RMean = zoo::rollmean(x = Distribution, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat) %>% 
  filter(
    # spp =="Centropristis striata"
    spp %in% quota_allocation$spp
  )

head(quota_df) %>% 
  kable()
```

### Exploring different ways to show the results

#### Difference line plot

This version shows the difference between the Historic quota allocation and the distribution proportion of the stock:

- diff < 0 = the distribution proportion is *larger* than the allocated quota
- diff > 0 = the distribution proportion is *smaller* than the allocated quota

```{r}
# Option one

quota_df %>%
  mutate(diff = Historic- Distribution,
         diff_rmean = RMean- Distribution) %>% 
  gather("type","value",diff:diff_rmean) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = year,
      y = value,
      color = order
    )
  ) +
  scale_color_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  facet_grid(type~spp, 
             scales = "free") +
  theme_dark()

```
#### Linetype plot

Here, we plot the distributional quota (solid lines) with each State's allocation with dashed (----) lines

```{r}
# Option one

quota_df %>% 
  gather("level","quota",Distribution:RMean) %>% 
  # mutate(diff = quota_per- percentage) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = year,
      y = quota,
      color = order
    )
  ) +
  labs(x = "Year",y = "Quota (%)") +
  scale_fill_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  scale_color_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  scale_linetype("Quota") +
  facet_grid(spp ~ level)+
  theme_dark()

```
#### Efficiency index

The idea here is create and *efficiency index* (Danger!) that computes the alignment between the distribution and the actual allocation. The index is simply;

$$ei = \frac{HistoricQuota}{DistributionProportion}$$ 
So, in cases where *ei* = 1, the quota is aligned with the distribution; when *ei* > 1 then the historic quota is more than the stock's State's distribution, contrarily, *ei* < 1 means that the historic quota is less than what the State currently has.

**Note:** 
- There are some crazzy outliers that have been removed, for now... 
- Bottom plot representing index as a Rmean

##### As a normal year to year

```{r}
# Option one

eff_index <- quota_df %>% 
  # filter(
  #   spp =="Centropristis striata"
  # ) %>% 
  mutate(index = Distribution/Historic,
         index_rmean = RMean/Historic) %>%
  gather("level","index",index:index_rmean) %>%
  filter(index < 5) # there are some craaaazy outliers
  # View()

ggplot(eff_index) +
  geom_line(
    aes(
      x = year,
      y = index,
      color = state
    ),
  ) +
  labs(x = "Year",y = "Efficiency index") +
  scale_fill_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  scale_color_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  facet_grid(level~spp,
             scales = "free")+
  theme_dark()

```

## NOAAs port approach

```{r noaa_landings, eval = F, echo = F}

noaa_landings <- read.csv("/Volumes/Enterprise/Data/AcrossBoundaries/Data/NOAA/landings_by_top_us_ports.csv")

noaa_landings %>% 
  clean_names() %>% 
  View()

state_totals <- as.data.frame(str_split_fixed(noaa_landings$Port., ",",2)) %>% 
  rename(name = V1,
         abrv = V2) %>% 
  bind_cols(noaa_landings) %>% 
  mutate(abrv = str_squish(abrv)) %>% 
  clean_names() %>% 
  filter(abrv %in%  c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")) %>% 
  filter(year >= 2010) %>% 
  group_by(abrv) %>% 
  summarise(total_pounds_million = sum(millions_of_pounds),
            total_dlr_million = sum(x_millions_of_dollars)
  )

# Estimate percentages
port_data <- as.data.frame(str_split_fixed(noaa_landings$Port., ",",2)) %>% 
  rename(name = V1,
         abrv = V2) %>% 
  bind_cols(noaa_landings) %>% 
  mutate(abrv = str_squish(abrv)) %>% 
  clean_names() %>% 
  filter(abrv %in%  c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")) %>% 
  filter(year >= 2010) %>% 
  group_by(abrv,name,port) %>% 
  summarise(sum_pounds_million = sum(millions_of_pounds),
            sum_dlr_million = sum(x_millions_of_dollars)
            ) %>% 
  left_join(state_totals) %>% 
  mutate(
    per_sum = sum_dlr_million/total_dlr_million*100,
    per_pnds = sum_pounds_million/total_pounds_million*100,
    cum_per_sum = cumsum(sum_dlr_million/total_dlr_million*100),
    cum_per_pnds = cumsum(sum_pounds_million/total_pounds_million*100)
  )

  # Function to get percentiles
      # https://www.tidyverse.org/blog/2020/03/dplyr-1-0-0-summarise/
      quibble <- function(x, q = p) {
        tibble(x = quantile(x, q), q = q)
      }
      # Get the percentile treshold per year, ensemble and neighbour id
    percentile <- port_data %>% 
        group_by(abrv) %>%
        do(quibble(.$com_sum, 0.50)) %>% 
      left_join(port_data) %>% 
      mutate(drop = ifelse(x < sum_dlr_million,"drop","keep" )) %>% 
      filter(drop == "keep")

# Number of ports per state
percentile %>%
  group_by(abrv) %>%
  summarise(n()) %>% 
  left_join(original_n) %>% View()

unique(port_data$abrv)
length(unique(port_data$abrv))

# Get rank by dolar
top_dlr <- port_data %>% 
  group_by(abrv) %>%
  # View()
  top_n(10,sum_dlr_million)

# Get rank by landing
top_pnd <- port_data %>% 
  group_by(abrv) %>%
  # View()
  top_n(10,sum_pounds_million)

# Bind ports and manually set corrdinates
top_ports <- top_dlr %>% 
  full_join(top_pnd) %>% 
  arrange(abrv) %>% 
  group_by(abrv) %>% 
  summarise(n())

  # Manually estimated from google maps
  bind_cols(
    lat = c(41.3345476816744,
            41.34138388600569,
            41.636467065626505,
            38.32347787720262, 
            44.15386541519202, 
            43.66617357876293,
            35.90400984508054, 
            43.1123838312889,
            38.971394018750296, 
            41.05268318179009, 
            41.37054800823896, 
            36.992858601329836, 
            37.839637137249234
            ),
    long = c(-71.91332744028196,
             -72.08202556545858,
             -70.91964446265216,
             -75.1014183484163,
             -68.66402196775981,
             -70.2288380196981,
             -75.55554059613718,
             -70.85815866495484,
             -74.8122413826507,
             -71.9688254628936,
             -71.48032452726083,
             -76.26817146730399,
              -76.24292771914955
             )
  ) %>% 
  # Delaware
  bind_rows(
    data.frame(abrv = "DE",
      name = "Indian river",
      port = "Indian river, DE",
      lat = as.double(38.62113593403746),
      long = -75.06397021718345
      )
  )

# write_csv(top_ports,"top_neus_ports.csv")

  port_data %>% 
    filter(com_mean >= 75) %>% 
    group_by(abrv) %>% 
    View()
    summarise(n())

```

```{r remove_pen, eval = F, echo = F}

####____________________________________________### 
# Solving for Issue #8 Pennsylvania has NO coast
####____________________________________________###

# Remove Pennsylvania from sf
grid_eez_sf <-  st_read("/Volumes/Enterprise/Data/AcrossBoundaries/Data/Spatial/grid_sf/grid_eez_sf.shp") %>%
  filter(state != "pennsylvania")

# Make sure no Penn.
unique(grid_eez_sf$state)

# over write sf
write_sf(grid_eez_sf, paste0(my_path("D", "Spatial"),"grid_eez_sf.shp"))

#Make sure of things

# grid_eez_sf <-  st_read("/Volumes/Enterprise/Data/AcrossBoundaries/Data/Spatial/grid_sf/grid_eez_sf.shp")

# unique(grid_eez_sf$state)

# Now save it as a DF too

# grid_eez_df <- as.data.frame(grid_eez_sf) %>%
  # select(state,index) %>%
  # left_join(ne_grid,
            # by = "index")

# write_csv(grid_eez_df, paste0(my_path("R", "Partial"),"grid_eez_df.csv"))
```

### Dynamic figure: Aggregated Proportional change per survey

```{r area_map_aggregated, eval = T, echo = T, fig.width = 10, fig.height=6}

total_fited <- historic_spp %>% 
  group_by(year,region,spatial) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_fit <- historic_spp %>% 
  group_by(state,year,region,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  left_join(periods,
            by = "year") %>% 
  group_by(state,order,label,region,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop")

# check percentages
# state_fit %>% 
#   group_by(period) %>% 
#   summarise(sum(mean_per))

p <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  # geom_sf(aes(fill = mean_per)) +
  geom_sf(aes(fill = mean_per, text = paste(state, mean_per,"% of proportion"))) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ region + label +spatial, 
             nrow = 2) +
  labs(x = "Latitude", 
       y = "Longitude") +
  my_ggtheme_p(facet_tx_s = 12,
               leg_pos = "bottom") +
  theme(legend.key.width = unit(1,"line")) +
  theme_dark()

ggplotly(p,
         tooltip = "text",
         dynamicTicks = FALSE) %>% 
    style(hoverlabel = list(bgcolor = "white"), hoveron = "fill")

```

```{r eval = F, echo = T, results='hide'}

total_fited <- historic_spp %>%
  group_by(year,region,spp,spatial) %>%
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_change <-historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  # Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)
         ) %>% 
  select(-region) %>% 
  spread(label,mean_per) %>% 
  mutate(change = (Today-Reference)/(Today)*100,
         change = ifelse(change > 100,100,
                         ifelse(change < -100,-100,change)
         )
  ) %>%
  select(-4,-5) %>% 
  spread(spatial,change) %>% 
  mutate(
    #difference = (fp-sw)/((fp+sw)/2)*100,
    difference = abs(fp-sw)#,
         # difference = ifelse(difference > 100,100,
         #                     ifelse(difference < -100,-100,difference)
         # )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         ),
         mean_per= replace_na(mean_per,0)# for those 0/0
  )

# The plot
prop_chng_map <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Change in proportion today\nrelative to 1980-2001\n(%)", alpha = 0.8,
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c("<-100",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5)

# ggsave(filename = "temp_proportion_chng.jpg",
#          plot = prop_chng_map,
#          path = my_path("R","Figures"),
#          width = 10,
#          height = 10
#   )

#### Option with differences

prop_chng_map_diff <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Change in proportion today\nrelative to 1980-2001\n(%)", alpha = 0.8,
                              option = "C"
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c("<-100",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5,
        strip.background = element_blank(),
        strip.text.x = element_blank())

# Cowplot option
p <- ggdraw() +
  # Revenue circular
  draw_plot(prop_chng_map, x = 0, y = 0.25, width = 1, height = 0.65) +
  # Catch circular
  draw_plot(prop_chng_map_diff, x = 0.101, y = 0, width = 0.798, height = 0.4) +
  draw_plot_label(label = c("Latitude", "Longitude"), 
                  size = 16,
                  angle = c(90,0),
                  x = c(0.1,0.45), 
                  y = c(0.45,0.11)
  )

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_diffN.jpg",
  p,
  base_height = 14,
  base_width = 14
)

```

### Proportion change realtive to past

```{r paper_figure_area_map, eval = F, echo = T, results='hide'}

total_fited <- historic_spp %>%
  group_by(year,season,spp,spatial) %>%
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_change <-historic_spp %>% 
  group_by(state,year,season,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  # Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)
         ) %>% 
  select(-region) %>% 
  spread(label,mean_per) %>% 
  mutate(change = (Today-Reference)/(Today)*100,
         change = ifelse(change > 100,100,
                         ifelse(change < -100,-100,change)
         )
  ) %>%
  select(-4,-5) %>% 
  spread(spatial,change) %>% 
  mutate(
    #difference = (fp-sw)/((fp+sw)/2)*100,
    difference = abs(fp-sw)#,
         # difference = ifelse(difference > 100,100,
         #                     ifelse(difference < -100,-100,difference)
         # )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         ),
         mean_per= replace_na(mean_per,0)# for those 0/0
  )

# The plot
prop_chng_map <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Change in proportion today\nrelative to 1980-2001\n(%)", alpha = 0.8,
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c("<-100",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5)

# ggsave(filename = "temp_proportion_chng.jpg",
#          plot = prop_chng_map,
#          path = my_path("R","Figures"),
#          width = 10,
#          height = 10
#   )

#### Option with differences

prop_chng_map_diff <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Change in proportion today\nrelative to 1980-2001\n(%)", alpha = 0.8,
                              option = "C"
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c("<-100",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5,
        strip.background = element_blank(),
        strip.text.x = element_blank())

# Cowplot option
p <- ggdraw() +
  # Revenue circular
  draw_plot(prop_chng_map, x = 0, y = 0.25, width = 1, height = 0.65) +
  # Catch circular
  draw_plot(prop_chng_map_diff, x = 0.101, y = 0, width = 0.798, height = 0.4) +
  draw_plot_label(label = c("Latitude", "Longitude"), 
                  size = 16,
                  angle = c(90,0),
                  x = c(0.1,0.45), 
                  y = c(0.45,0.11)
  )

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_diffN.jpg",
  p,
  base_height = 14,
  base_width = 14
)

```

### Aggregated Proportional change

```{r area_map_aggregated_formans, eval = T, echo = T, fig.width = 10, fig.height=6}

total_fited <- historic_spp %>%
  group_by(year,region,spatial) %>%
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

state_change <- historic_spp %>% 
  group_by(state,year,region,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"reference",
                         ifelse(year > 2001,"today",NA)
                         )
         ) %>% 
  # left_join(periods,
            # by = "year") %>% 
  group_by(state,label,region,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(#str_detect(region,"Spring"),
         !is.na(label)
         ) %>% 
  # select(-region) %>% 
  # Estimate percentage change
  spread(label,mean_per) %>% 
  mutate(raw_change = (today-reference)/(today)*100,
         change = ifelse(raw_change > 100,100,
                         ifelse(raw_change < -100,-100,raw_change)
         )
  ) %>%
  select(-reference,-today)

#### Average
state_change %>% 
  mutate(
    wl = ifelse(change < 0, "L",
                ifelse(change == 0, "NC","W"))
  ) %>% 
  # View()
  # group_by(wl) %>% 
  group_by(region,spatial,wl) %>% 
  summarise(
    states = paste(unique(state),collapse = ", "),
    mean(raw_change),
    sd(raw_change)
  ) %>% 
  write_csv("../Results/state_relative_chng.csv")
  View()

  # Estimate spatial differences in percentage change
spatial_diff <- state_change %>% 
  spread(spatial,change) %>% 
  mutate(spatial_diff = abs(fp-sw)) %>% 
  select(state, category = region, diff = spatial_diff)

# Estimate seasonal differences in percentage change
season_diff <- state_change %>% 
  spread(region,change) %>%  
  mutate(season_diff = abs(`Northeast US Fall`-`Northeast US Spring`)) %>% 
  select(state,category = spatial,diff = season_diff)

differences <- season_diff %>% 
  bind_rows(spatial_diff) %>% 
  mutate(
    label = ifelse(category == "fp","Fishing ports",
                          ifelse(category == "sw","State Waters",category)
         )
  )

# For in text differences
differences %>% 
  group_by(category) %>%
  summarise_at(vars(diff),
    c(mean = mean,
      sd =sd,
      min = min,
      max= max))

# The main plot
prop_chng_map <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  # filter(spatial != "Difference" & cat != "difference_reg") %>% 
  ggplot() +
  geom_sf(aes(fill = change)) +
  viridis::scale_fill_viridis("Change in proportion today\n relative to 1973-1984\n(%)", alpha = 0.8,
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c(seq(-100,75,25),">100")
                              ) +
  facet_grid(spatial~ region) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5,
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        )

ggsave(filename = "proportion_chng_agg.png",
         plot = prop_chng_map,
         path = my_path("R","Figures"),
         width = 10,
         height = 10
  )

#### Option with differences

# Season differencce

spat_diff <- land_sf %>% 
  left_join(differences,
            by = "state") %>% 
  filter(!category %in% c("fp","sw")) %>% 
  ggplot() +
  geom_sf(aes(fill = diff)) +
  viridis::scale_fill_viridis("Differences", 
                              alpha = 0.8,
                              option = "C",
                              limits = c(0,50),
                              breaks = seq(0,50,10)
                              ) +
  facet_wrap(~ label,
             ncol = 2) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5,
        strip.background = element_blank(),
        strip.text.x = element_blank())

ggsave(filename = "spat_diff.png",
         plot = spat_diff,
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )

# seasonal
seas_diff <- land_sf %>% 
  left_join(differences,
            by = "state") %>% 
  filter(category %in% c("fp","sw")) %>% 
  ggplot() +
  geom_sf(aes(fill = diff)) +
  viridis::scale_fill_viridis("Differences", 
                              alpha = 0.8,
                              option = "C",
                              limits = c(0,50),
                              breaks = seq(0,50,10)
                              ) +
  facet_wrap(~ label,
             ncol = 1,
             strip.position="right") +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(strip.background = element_blank(),
        # strip.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank()
        )

ggsave(filename = "seas_diff.png",
         plot = seas_diff,
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )

```

```{r}

### SRAH LANDING DATA
# read_excel_allsheets <- function(filename, tibble = TRUE) {
#     # I prefer straight data.frames
#     # but if you like tidyverse tibbles (the default with read_excel)
#     # then just pass tibble = TRUE
#     sheets <- readxl::excel_sheets(filename)[2:8]
#     x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
#     if(!tibble) x <- lapply(x, as.data.frame)
#     names(x) <- sheets
#     x
# }
# Get data path
# ports_path <- my_path("D","NOAA", name = "SSmith_2015-2021 Landings_JUN 2021.xlsx")

# Load ports
# ports_data <- bind_rows(read_excel_allsheets(ports_path)) %>%
#   clean_names() %>%
#   mutate_if(is.character,
#               ~ str_to_lower(.)) %>%
#     filter(species_name %in% c("summer flounder","black sea bass","scup"))

```

*Note:* This plot is interactive. For ease comparison between States you can;

- *One* click on any State to *remove* it from the plot 
- *Two* clicks on any State to *isolate it* from the plot (other states can then be added by just clicking on them).
- The bottom panel allows you to reduce the time frame

### Dynamic aggregated plot

```{r dynamic_area_plot, eval = T, echo = T, fig.height = 10, fig.width = 10}

total_fited <- historic_spp %>% 
  group_by(year,region) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100)

# Plot me
p <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(percentage),
      fill = state
    )
  ) +
  ylab("Percentage (%)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(values = state_pallet) +
  MyFunctions::my_ggtheme_p() +
  facet_wrap(~region, ncol = 1) +
  theme_dark()

ggplotly(p,
         dynamicTicks = TRUE) %>% 
  layout(hovermode = "x") %>%
  rangeslider()

```

### Dynamic aggregated plot

*Note:* This plot is also interactive and thus, follows the same options of the previous one.

```{r area_plot, eval = T, echo = T, fig.height = 10, fig.width = 10}

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,.groups = "drop") %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>%  
  left_join(state_lat)

# Plot me
p <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = state
      # fill = order
    )
  ) +
  ylab("10 yrs running average (%)") +
  scale_fill_manual(
    "State",
    values = state_pallet,
    # labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  MyFunctions::my_ggtheme_p() +
  facet_wrap(~region, ncol = 1) +
  theme_dark()

suppressWarnings(
ggplotly(p,
         dynamicTicks = TRUE) %>% 
  layout(hovermode = "x") %>% 
  # add_trace() %>% 
  rangeslider()
)

```

#### Aggregated Area plot for all species

```{r plot_area_ave_agg, eval = F, echo = F, results='hide'}

total_fited <- historic_spp %>% 
  group_by(year,region,spatial) %>% 
  summarise(total_value = sum(value,na.rm=T),.groups = "drop")

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                            10,
                            fill = NA,
                            na.rm=T)
    ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
         ) %>%  
  left_join(state_lat) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         )
  )

# Plot me
agg_area <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  annotate(geom= "rect",
           xmin = 1980, xmax = 2001, 
           ymin = 0, ymax = 100,
            fill = "gray90",
           colour = NA,
           alpha = 0.5) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State\n(Latitudinal order)",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 1,ax_tx_s = 10,leg_tx_s = 10, leg_tl_s = 12,ax_tl_s = 12) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5) +
  facet_wrap(~spatial)

    ggsave(filename = "area_plot_avg_agg_sptl.jpg",
         plot = agg_area,
         path = my_path("R","Figures"),
         width = 8,
         height = 6
         )

```