### 2022-08-09 Scripts/nefsc_index_script.Rmd
Estimated grid area for Chris D.

---
title: "Merge our grid with NEFSC"
output: html_document
---

# Instructions

Would it be possible for you to add a column that gives the NOAA-NMFS Statistical Area number for each row in the data file?

# Data

## NEFSC Shapefile

Shapefile provided by Chris on the NEFSC website:  ftp://ftp.nefsc.noaa.gov/pub/gis/

- Statistical_Areas_2010_withNames.shp
- Statistical_Areas_2010.shp

- The description and metadata for the shape file are here: https://www.fisheries.noaa.gov/inport/item/26262

This is the NEFSC-recommended citation for the shapefile:

Northeast Fisheries Science Center, 2021: NEFSC Statistical Areas from 2010-06-15 to 2010-08-15. NOAA National Centers for Environmental Information, https://www.fisheries.noaa.gov/inport/item/26262.

## Load Data

```{r load_data}

# oad our grid SF
grid_sf <- st_read(my_path("D","Partial/grid_sw", name = "grid_eez_sw_sf.shp")) %>% 
  st_transform(4326) %>% # to match shape
  st_set_crs(4326) %>% 
  select(index,geometry)

#Load NEFSC shapefile
nefsc_sf <- st_read(my_path("D","Spatial/nefsc", name = "Statistical_Areas_2010.shp")) %>% 
  st_transform(4326) %>% # to match shape
  st_set_crs(4326) %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)

# Visualize map (ok)
# nefsc_sf %>% 
#   st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
#   ggplot() +
  # geom_sf()

```

## Merge shapefiles

```{r merge_shapefiles}

# Merge both sf
nefsc_index <- nefsc_sf %>% 
   st_join(grid_sf,
           join = st_intersects) %>% 
   filter(!is.na(index)) %>% 
  as.data.frame() %>% 
  select(Id,index) %>% 
  distinct(.keep_all = T) %>% 
  mutate(plop = "yes") %>% 
  left_join(my_path("D", "Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T)) %>% 
  select(Id,index,state)

# Make sure it works
nefsc_sf %>%
  left_join(nefsc_index) %>%
  filter(!is.na(index)) %>%
  ggplot() +
  geom_sf(aes(fill = index))

```

## Save shapefile

```{r}

# File name
file_name <- my_path("D","Spatial", name = "nefsc_index.csv")

write_csv(nefsc_index %>% select(-state), file_name)
```

# Estimate area of gridcell

Would it be possible for you to send me a file with the area in hectares for each of the “index” grid areas in your fish distribution analysis?  You sent me kg/ha as one field in the dataset, so I’m hoping that you also have hectares for each index area, and that it’s easy to send.

```{r area_estimation}

library("tidyverse")

# Load coordinate grid
coords <- MyFunctions::my_path("D", extra_path = "Spatial", name = "inter_grid_df.csv", read = T, header = T)

# 1.2 Estimate gridcell area based on global variables (m^2)
# Degree ressolution for area estimation
# https://rdrr.io/github/stineb/rbeni/src/R/area.R
# https://www.pmel.noaa.gov/maillists/tmap/ferret_users/fu_2004/msg00023.html
# https://www.engr.scu.edu/~emaurer/tools/calc_cell_area_cgi.pl

r_earth <- 6371000 # NASA's volumetric mean radious https://nssdc.gsfc.nasa.gov/planetary/factsheet/earthfact.html
res_lon <- 0.3 # lon degree resolution
res_lat <- 0.3 # lat degree resolution

coords <- coords %>% 
  mutate(
    area = (4 * r_earth^2 * 0.5 * res_lon * pi/180 * cos(abs(lat) * pi/180 ) * sin(0.5 * res_lat * pi/180))/10000
  )

write_csv(coords, "corrds_area.csv")

```

```{r plot_test}

ggplot(coords) +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = area
    )
  ) +
  scale_fill_gradient(low = "yellow", high = "red", na.value = NA)

```